---
title: "Spatial Priorization Model Results for Scenario 4 & 5"
author: Anna Abelman, GP-Future4Fins
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = F, warning = F, message = F)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(rgdal)
library(ggmap)
library(tmap)
```

```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
#plot(exist_mpas)
```


```{r}
## create function
create_map <- function(final_raster, mpas){

  final_raster[final_raster == 0] <- NA
  
  tmap_mode("view") # Set to interactive viewing

  tm_shape(final_raster)+
    tm_raster(final_raster[1], palette="dodgerblue3", legend.show = F)+
  tm_shape(mz_eez_sf)+
    tm_borders("black", lwd = .5) +
  tm_basemap("Esri.WorldGrayCanvas")+
    tm_shape(exist_mpas)+
  tm_raster(exist_mpas[1], palette = "deeppink2", legend.show = F)+
    tm_add_legend("fill",
              col = c("dodgerblue3", "deeppink2"),
	            border.col = "black",
	            labels = c('Suggested Areas','Existing MPAs'),
	            title="Prioritization Model")
}
```

Future4Fins GP team developed a spatial prioritization model that identifies key areas for shark and ray conservation within the Mozambique Economic Exclusive Zone (EEZ) for our client, Wildlife Conservation Society. This model will be used by the Mozambique government to inform a greater MPA network across the region. There were five different scenarios, each with a different objective and a unique cost layer. Below shows Scenario 4 (Minimal Impact on all Fisheries) and Scenario 5 (Target Areas with High Fishing Pressure).


#### **Scenario 4: Minimal Impact on all Fisheries**
```{r}
### create file path to read in final raster
combined_path <- here("blm_sensitivity")
combined_files <- list.files(combined_path, full.names = TRUE)

combined_20_blm0.000001 <- raster(combined_files[4])
#plot(combined_20_blm0.000001)


## create interactive maps for combined sc
combined_20_blm0.000001_map <- create_map(combined_20_blm0.000001, exist_mpas)
combined_20_blm0.000001_map
```

**Figure 1**: Prioritizr results for Scenario 4 - Minimal Impact on all Fisheries: using combined fishing pressure as cost, boundary penalties of 10x6-6 with conservation targets of 20%, and existing marine protected areas locked-in. 


#### **Scenario 5: Target Areas with High Fishing Pressure**
```{r}
### create file path to read in final raster
inverse_path <- here("final_results", "scenario_5_inverse")
inverse_files <- list.files(inverse_path, full.names = TRUE)

inverse_20_blm0.000001 <- raster(inverse_files[3])
#plot(inverse_20_blm0.000001)

## create interactive maps for inverse sc
inverse_20_blm0.000001_map <- create_map(inverse_20_blm0.000001, exist_mpas)
inverse_20_blm0.000001_map


```

**Figure 2**: Prioritizr results for Scenario 5 - Target Areas with High Fishing Pressure: using the inverse of combined fishing pressure as cost, boundary penalties of 10x6-6 with conservation targets of 20%, and existing marine protected areas locked-in.

Data from: IUCN, Allen Coral Altas, WCS, VMS data from the Mozambique government

