---
title: "Scenario 4 - Combined Fishing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)
```

## Scenario 2 - Combined Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on both artisanal and indsutrial fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_4 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

# ### plot to make sure it looks okay
plot(pu_4)
```


#### Cost Layer
```{r}
### pull in the combined fishing pressure tif file from costs folder in repo
cost_combined <- raster(here("set_up", "final_rasters", "costs", "combined_stack_sum.tif"))

### plot to make sure it looks okay
plot(cost_combined)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack_4 <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack_4)
```

### Additional Files (Locked-in & Locked-out)
```{r}
# MPAs
### read in existing mpas raster
exist_mpas_4 <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas_4)
```


```{r}
# Key Biodiversity Areas (KBAs)
### read in existing kbas raster
exist_kbas <- raster(here("set_up", "final_rasters", "planning_unit", "kba.tif"))

### plot to make sure it looks okay
plot(exist_kbas)
```


```{r}
# Aggregation Sites
### read in existing aggregation sites raster
exist_aggs <- raster(here("set_up", "final_rasters", "planning_unit", "aggregation_sites.tif"))

### plot to make sure it looks okay
plot(exist_aggs)
```

## **STEP 2: Run Some combined Fishing Scenarios with no boundary penalty**

**Changing the targets!**

#### RUN 1: combined - 10% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 10% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_10 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_combined_10 <- solve(prob_combined_10)

## Plot the solution to see what it looks like
plot(sprob_combined_10,  main = c("combined- 10% Targets"))

#Save plot
writeRaster(sprob_combined_10, here("final_results", "scenario_4_combined", "sprob_combined_10_6hr_blm0.tif"), overwrite = TRUE)

```


#### RUN 2: combined - 20% target, Boundary Penalty = 0,  
```{r}

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) 

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)

## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

## Save plot
writeRaster(sprob_combined_20, here("final_results", "scenario_4_combined", "sprob_combined_20_6hr_blm0.tif"), overwrite = TRUE)

```

#### RUN 3: combined - 30% target, Boundary Penalty = 0,  
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_30 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) 

## Solve problem
sprob_combined_30 <- solve(prob_combined_30)

## Plot the solution to preview
plot(sprob_combined_30,  main = c("combined- 30% Targets"))

## Save plot
writeRaster(sprob_combined_30, here("final_results", "scenario_4_combined", "sprob_combined_30_6hr_blm0.tif"), overwrite = TRUE)

```


#### RUN 4: combined - 50% target, Boundary Penalty = 0,  
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_50 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) 

## Solve problem
sprob_combined_50 <- solve(prob_combined_50)

## Plot the solution to preview
plot(sprob_combined_50,  main = c("combined- 50% Targets"))

## Save plot
writeRaster(sprob_combined_50, here("final_results", "scenario_4_combined", "sprob_combined_50_6hr_blm0.tif"), overwrite = TRUE)

```



***

**Run some Industrial Fishing Scenarios with efficient boundary penalty (1x10^-6)**

#### Run 5: Industrial - 10% target, Boundar Penaly = 0.000001
```{r}
### OBJECTIVE: Minimize industrial fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_10_blm <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_combined_10_blm <- solve(prob_combined_10_blm)

## Plot the solution to see what it looks like
plot(sprob_combined_10_blm,  main = c("combined - 10% Targets"))

#Save plot
writeRaster(sprob_combined_10_blm, here("final_results", "scenario_4_combined", "sprob_combined_10_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### RUN 6: Industrial - 20% target, Boundary Penalty = 0.000001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20_blm <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_combined_20_blm <- solve(prob_combined_20_blm)

## Plot the solution to see what it looks like
plot(sprob_combined_20_blm,  main = c("combined - 20% Targets"))

#Save plot
writeRaster(sprob_combined_20_blm, here("final_results", "scenario_4_combined", "sprob_combined_20_6hr_blm0.000001.tif"), overwrite = TRUE)
```

#### RUN 7: Industrial - 30% target, Boundary Penalty = 0.000001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_30_blm <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_combined_30_blm <- solve(prob_combined_30_blm)

## Plot the solution to preview
plot(sprob_combined_30_blm,  main = c("combined- 30% Targets"))

#Save plot
writeRaster(sprob_combined_30_blm, here("final_results", "scenario_4_combined", "sprob_combined_30_6hr_blm0.000001.tif"), overwrite = TRUE)
```

#### RUN 8: Industrial - 50% target, Boundary Penalty = 0.000001 
```{r}
### OBJECTIVE: Minimize industrial fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_50_blm <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_combined_50_blm <- solve(prob_combined_50_blm)

## Plot the solution to preview
plot(sprob_combined_50_blm,  main = c("combined- 50% Targets"))

## Save plot
writeRaster(sprob_combined_50_blm, here("final_results", "scenario_4_combined", "sprob_combined_50_6hr_blm0.000001.tif"), overwrite = TRUE)
```

Other sensitivity analysis:

#### Run 9: Combined - 20% target, boundary = 0 and no MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_20_nompa <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_combined_20_nompa <- solve(prob_combined_20_nompa)

## Plot the solution to see what it looks like
plot(sprob_combined_20_nompa,  main = c("combined- 20% Targets with no mpas"))

## Save plot
writeRaster(sprob_combined_20_nompa, here("final_results", "sensitivity", "mpas", "sprob_combined_20_6hr_blm0_nompa.tif"), overwrite = TRUE)
```

#### Run 10: Combined - 20% target, boundary = 0.000001 and no MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_20_blm_nompa <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_combined_20_blm_nompa <- solve(prob_combined_20_blm_nompa)

## Plot the solution to see what it looks like
plot(sprob_combined_20_blm_nompa,  main = c("combined- 20% Targets with boundary penalty and no mpas"))

## Save plot
writeRaster(sprob_combined_20_blm_nompa, here("final_results", "sensitivity", "mpas", "sprob_combined_20_6hr_blm0.000001_nompa.tif"), overwrite = TRUE)
```


#### Run 10: Combined - 20% target, boundary = 0 and KBAs & MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>% 
  add_locked_in_constraints(exist_kbas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0)

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20, here("final_results", "sensitivity", "key_biodiversity_areas", "sprob_combined_20_6hr_blm0_mpas_kbas.tif"), overwrite = TRUE)
```


#### Run 11: Combined - 20% target, boundary = 0 and KBAs & MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  #add_locked_in_constraints(exist_mpas_4) %>% 
  #add_locked_in_constraints(exist_kbas) %>%
  #add_locked_in_constraints(exist_aggs) %>% 
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0)

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20, here("final_results", "sensitivity", "sprob_combined_20_6hr_blm0_aggsites.tif"), overwrite = TRUE)

