---
title: "scenario_6_management"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)
```

# Scenario 6 - Management 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while remaining relatively close to the shore for management purposes (<50m depth).  For this run, we used the inverse of the combined fishing distance decay raster as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

Step 1 reads in all the necessary data
Step 2 run to see results of varying targets from 10-50% and boundary penalties

## **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
### reading in the baseline MZ EEZ raster (with cell ids) 
pu_6 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

### plot to make sure it looks okay
plot(pu_6)
```

#### Cost Layer
```{r}
### pull in the cost layer created in setup step 9
cost_distance <- raster(here("set_up", "final_rasters", "costs", "dist_rast.tif"))

### plot to make sure it looks okay
plot(cost_distance)
```

#### Conservation Features
This includes all species distributions and habitat rasters:
```{r}
### pull in species stack that we created in set-up step 3
features_stack_6 <- brick(here("set_up", "final_rasters", "conservation_features","all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack_6)
```

#### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas)
```


#### Set the some of the global model parameters like gap, time limit, and efficient boundary penalty to use in soem runs:
```{r}
# if you decide to change the gap, time-limit or boundary penalty value, change it here and rerun the subsequent model runs
gap <- 0.01
time_limit <- 21600
bp <- 0.000001
```

## **STEP 2: Run Models with varying targets**

#### RUN 1: Baseline - 10% target, Boundary Penalty = 0 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_management_10 <- problem(cost_distance, features = features_stack_6) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_management_10 <- solve(prob_management_10)

## Plot the solution to see what it looks like
plot(sprob_management_10,  main = c("Area- 10% Targets"))

## Save as tif
writeRaster(sprob_management_10, here("final_results", "scenario_6_management",  "sprob_management_10_6hr_blm0.tif"), overwrite = TRUE)
```

#### RUN 2: Baseline - 10% target, Boundary Penalty = 0.00001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_management_10_blm <- problem(cost_distance, features = features_stack_6) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(bp)

## Solve problem
sprob_management_10_blm <- solve(prob_management_10_blm)

## Plot the solution to see what it looks like
plot(sprob_management_10_blm,  main = c("Area- 10% Targets"))

## Save as tif
writeRaster(sprob_management_10_blm, here("final_results", "scenario_6_management",  "sprob_management_10_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### RUN 3: Baseline - 5% target, Boundary Penalty = 0 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_management_.05 <- problem(cost_distance, features = features_stack_6) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.05) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_management_.05 <- solve(prob_management_.05)

## Plot the solution to see what it looks like
plot(sprob_management_.05,  main = c("Area- 5% Targets"))

## Save as tif
writeRaster(sprob_management_.05, here("final_results", "scenario_6_management",  "sprob_management_5_6hr_blm0.tif"), overwrite = TRUE)
```

#### RUN 4: Baseline - 5% target, Boundary Penalty = 0.00001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_management_.05_blm <- problem(cost_distance, features = features_stack_6) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.05) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(bp)

## Solve problem
sprob_management_.05_blm <- solve(prob_management_.05_blm)

## Plot the solution to see what it looks like
plot(sprob_management_.05_blm,  main = c("Area- 5% Targets"))

## Save as tif
writeRaster(sprob_management_10_blm, here("final_results", "scenario_6_management",  "sprob_management_5_6hr_blm0.000001.tif"), overwrite = TRUE)
```

