---
title: "Scenario 4 - Combined Fishing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)

```

## Scenario 2 - Combined Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on both artisanal and indsutrial fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_4 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

# ### plot to make sure it looks okay
plot(pu_4)
```


#### Cost Layer
```{r}
### pull in the combined fishing pressure tif file from costs folder in repo
cost_combined <- raster(here("set_up", "final_rasters", "costs", "combined_stack_sum.tif"))

### plot to make sure it looks okay
plot(cost_combined)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack_4 <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack_4)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas_4 <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas_4)
```

## **STEP 2: Run Some combined Fishing Scenarios!**

**Changing the targets!**

#### SCENARIO 1: combined - 20% target, Boundary Penalty = 0,  
```{r}
### OBJECTIVE: Minimize combined fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) 

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))
```

#### SCENARIO 2: combined - 10% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 10% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_10 <- problem(cost_combined, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_combined_10 <- solve(prob_combined_10)
## Plot the solution to see what it looks like
plot(sprob_combined_10,  main = c("combined- 10% Targets"))
```


#### SCENARIO 3: combined - 50% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 50% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_combined_50 <- problem(cost_combined, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_combined_50 <- solve(prob_combined_50)
## Plot the solution to see what it looks like
plot(sprob_combined_50,  main = c("combined- 50% Targets"))
```



***

## BLM Sensitivity Analysis

##1 Royal Penguin
#### BLM: 0.01, Combined, 20% Target, MPAs locked in - 20% target, 6hr time limit
```{r}
### OBJECTIVE: Minimize combined fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0.0000001

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000003)

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20, here("final_results", "sprob_combined_20_6hr_blm0.000003.tif"), overwrite = TRUE)

```

##2 Islay
#### BLM: .01, Combined, 20% Target, MPAs locked in - 20% target, 6hr time limit
```{r}
### OBJECTIVE: Minimize combined fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0.0000001

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.01)

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20, here("final_results", "sprob_combined_20_6hr_blm0.01.tif"), overwrite = TRUE)

```

##3 Dugong
#### BLM: .0001, Combined, 20% Target, MPAs locked in - 20% target, 6hr time limit
```{r}
### OBJECTIVE: Minimize combined fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0.0000001

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000007)

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20, here("final_results", "sprob_combined_20_6hr_blm0.000007.tif"), overwrite = TRUE)

```

##4 Mason
#### BLM: .0001, Combined, 20% Target, MPAs locked in - 20% target, 6hr time limit
```{r}
### OBJECTIVE: Minimize combined fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0.0000001

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20 <- problem(cost_combined, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.0001)

## Solve problem
sprob_combined_20 <- solve(prob_combined_20)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20, (here("final_results", "sprob_combined_20_6hr_blm0.0001.tif")), overwrite = TRUE)

```

## Convert tifs to shapefiles to extract BLM data
```{r}
#BLM 0
blm0 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.tif"))
blm0_sf <- rasterToPolygons(blm0, fun = function(sprob_combined_20_6hr_blm0){sprob_combined_20_6hr_blm0==1})
plot(blm0_sf)
shapefile(blm0_sf, here("set_up", "blm_sensitivity", "blm0"))

#BLM 0.000000001
blm0.000000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000000001.tif"))
blm0.000000001_sf <- rasterToPolygons(blm0.000000001, fun = function(sprob_combined_20_6hr_blm0.000000001){sprob_combined_20_6hr_blm0.000000001==1})
plot(blm0.000000001_sf)
shapefile(blm0.000000001_sf, here("set_up", "blm_sensitivity", "blm81"))

#BLM 0.00000001
blm0.00000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.00000001.tif"))
blm0.00000001_sf <- rasterToPolygons(blm0.00000001, fun = function(sprob_combined_20_6hr_blm0.00000001){sprob_combined_20_6hr_blm0.00000001==1})
plot(blm0.00000001_sf)
shapefile(blm0.00000001_sf, here("set_up", "blm_sensitivity", "blm71"))

#BLM 0.0000001
blm0.0000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.0000001.tif"))
blm0.0000001_sf <- rasterToPolygons(blm0.0000001, fun = function(sprob_combined_20_6hr_blm0.0000001){sprob_combined_20_6hr_blm0.0000001==1})
plot(blm0.0000001_sf)
shapefile(blm0.0000001_sf, here("set_up", "blm_sensitivity", "blm61"))

#BLM 0.000001
blm0.000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000001.tif"))
blm0.000001_sf <- rasterToPolygons(blm0.000001, fun = function(sprob_combined_20_6hr_blm0.000001){sprob_combined_20_6hr_blm0.000001==1})
plot(blm0.000001_sf)
shapefile(blm0.000001_sf, here("set_up", "blm_sensitivity", "blm51"))

#BLM 0.00001
blm0.00001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.00001.tif"))
blm0.00001_sf <- rasterToPolygons(blm0.00001, fun = function(sprob_combined_20_6hr_blm0.00001){sprob_combined_20_6hr_blm0.00001==1})
plot(blm0.00001_sf)
shapefile(blm0.00001_sf, here("set_up", "blm_sensitivity", "blm41"))

#BLM 0.0001
blm0.0001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.0001.tif"))
blm0.0001_sf <- rasterToPolygons(blm0.0001, fun = function(sprob_combined_20_6hr_blm0.0001){sprob_combined_20_6hr_blm0.0001==1})
plot(blm0.0001_sf)
shapefile(blm0.0001_sf, here("set_up", "blm_sensitivity", "blm31"))

#BLM 0.001
blm0.001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.001.tif"))
blm0.001_sf <- rasterToPolygons(blm0.001, fun = function(sprob_combined_20_6hr_blm0.001){sprob_combined_20_6hr_blm0.001==1})
plot(blm0.001_sf)
shapefile(blm0.001_sf, here("set_up", "blm_sensitivity", "blm21"))

#BLM 0.01
blm0.01 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.01.tif"))
blm0.01_sf <- rasterToPolygons(blm0.01, fun = function(sprob_combined_20_6hr_blm0.01){sprob_combined_20_6hr_blm0.01==1})
plot(blm0.01_sf)
shapefile(blm0.01_sf, here("set_up", "blm_sensitivity", "blm11"))

#BLM 0.1
blm0.1 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.1.tif"))
blm0.1_sf <- rasterToPolygons(blm0.1, fun = function(sprob_combined_20_6hr_blm0.1){sprob_combined_20_6hr_blm0.1==1})
plot(blm0.1_sf)
shapefile(blm0.1_sf, here("set_up", "blm_sensitivity", "blm01"))

#BLM 1
blm1 <- raster(here("final_results", "sprob_combined_20_6hr_blm1.tif"))
blm1_sf <- rasterToPolygons(blm1, fun = function(sprob_combined_20_6hr_blm1){sprob_combined_20_6hr_blm1==1})
plot(blm1_sf)
shapefile(blm1_sf, here("set_up", "blm_sensitivity", "blm1"))

```

## BLM optimization curve plot
```{r}
#read in csv
blm_sensitivity <- read_csv(here("set_up", "blm_sensitivity", "blm_sensitivity_values.csv"))

# plot it - average area
ggplot(data = blm_sensitivity, aes(x = area_avg_km2, y = perimeter_sum_km)) +
  geom_point() +
  geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
  geom_line() +
  labs(x = "Average Area (km2)",
       y = "Boundary Length Sum (km)",
       title = "BLM Sensitivity Analysis")

# plot it - total area
ggplot(data = blm_sensitivity, aes(x = total_area_km2, y = perimeter_sum_km)) +
  geom_point() +
  geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
  geom_line() +
  labs(x = "Total Area (km2)",
       y = "Boundary Length Sum (km)",
       title = "BLM Sensitivity Analysis - Total Area")

# plot it - perimeter to area
ggplot(data = blm_sensitivity, aes(x = BLM, y = perimeter_to_area)) +
  geom_point() +
  geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
  geom_line() +
  labs(x = "Total Area (km2)",
       y = "Boundary Length Sum (km)",
       title = "BLM Sensitivity Analysis - Total Area")

```







