---
title: "Scenario 2: Artisanal Fishing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)

```

## Scenario 2 - Artisanal Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on artisanal fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_2 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))
# 
# ### plot to make sure it looks okay
plot(pu_2)
```


#### Cost Layer
```{r}
### pull in the artisanal fishing pressure tif file from costs folder in repo
cost_artisanal <- raster(here("set_up", "final_rasters", "costs", "cost_art_fish.tif"))

### plot to make sure it looks okay
plot(cost_artisanal)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas)
```

## **STEP 2: Run Some Artisanal Fishing Scenarios!**
1. Run Scenario 1 at different time limits (2min, 20min, 6hr, 24hr)
2. Changing the targets (20%, 10%, 50%) --> based time limit off step 1
3. Changing Boundary Penalty (0, 1, 0.001, 0.01) --> all 1 hour
4. Run without locking in existing mpas (BLM = 0, 20%) --> 1 hour
5. Run with locking in KBAs (BLM = 0, 20%) --> 1 hour

**Changing the targets!**

#### SCENARIO 1: Artisanal - 20% target, Boundary Penalty = 0.01,  
```{r}
### OBJECTIVE: Minimize artisanal fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0.01

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal_20 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_artisanal_20 <- solve(prob_artisanal_20)
## Plot the solution to see what it looks like
plot(sprob_artisanal_20,  main = c("Artisanal- 20% Targets"))

#Save plot

writeRaster(sprob_artisanal_20, here("final_results", "scenario_2_artisanal", "sprob_artisanal_20_6hr_blm0.000001.tif"), overwrite = TRUE)

```



#### SCENARIO 2: Artisanal - 10% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 10% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_10 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_artisanal_10 <- solve(prob_artisanal_10)
## Plot the solution to see what it looks like
plot(sprob_artisanal_10,  main = c("Artisanal- 10% Targets"))

#Save plot
writeRaster(sprob_artisanal_10, here("final_results", "scenario_2_artisanal", "sprob_artisanal_10_6hr_blm0.000001.tif"), overwrite = TRUE)

```


#### SCENARIO 3: Artisanal - 50% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 50% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_50 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_artisanal_50 <- solve(prob_artisanal_50)
## Plot the solution to see what it looks like
plot(sprob_artisanal_50,  main = c("Artisanal- 50% Targets"))

#Save plot
writeRaster(sprob_artisanal_50, here("final_results", "scenario_2_artisanal", "sprob_artisanal_50_6hr_blm0.000001.tif"), overwrite = TRUE)
```

**Changing BLM with constant target (20%)**

#### SCENARIO 4: Artisanal - 20% target, Boundary Penalty = 0  
```{r}
### OBJECTIVE: Minimize artisanal fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal_blm_0 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_boundary_penalties(penalty=0) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_artisanal_blm_0 <- solve(prob_artisanal_blm_0)
## Plot the solution to see what it looks like
plot(sprob_artisanal_blm_0,  main = c("Artisanal- 20% Targets"))

#Save plot
writeRaster(sprob_artisanal_20, here("final_results", "scenario_2_artisanal", "sprob_artisanal_20_6hr_blm0.tif"), overwrite = TRUE)

```

#### SCENARIO 5: Artisanal - 20% target, Boundary Penalty = 1 
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 1

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_blm_1 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_boundary_penalties(penalty=1) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_artisanal_blm_1 <- solve(prob_artisanal_blm_1)
## Plot the solution to see what it looks like
plot(sprob_artisanal_blm_1,  main = c("Artisanal- 20% Targets, BLM = 1"))
```


#### SCENARIO 6: Artisanal - 20% target, Boundary Penalty = 0.001 
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0.001

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_blm_0.001 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_boundary_penalties(penalty=0.001) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_artisanal_blm_0.001 <- solve(prob_artisanal_blm_0.001)
## Plot the solution to see what it looks like
plot(sprob_artisanal_blm_0.001,  main = c("Artisanal- 20% Targets, BLM = 0.001"))
```


#### SCENARIO 7: Artisanal - 20% target, Boundary Penalty = 0.01
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0.01

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_blm_0.01 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_boundary_penalties(penalty=0.01) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_artisanal_blm_0.01 <- solve(prob_artisanal_blm_0.01)
## Plot the solution to see what it looks like
plot(sprob_artisanal_blm_0.01,  main = c("Artisanal- 20% Targets, BLM = 0.01"))
```


**Remove MPAs**

#### SCENARIO 8: Artisanal - 20% target, Remove MPAs
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, without locking in existing MPAs and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_wo_mpa <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_artisanal_wo_mpa <- solve(prob_artisanal_wo_mpa)
## Plot the solution to see what it looks like
plot(sprob_artisanal_wo_mpa,  main = c("Artisanal- 20% Targets, No Existing MPAs"))
```

**Adding KBAs**

#### SCENARIO 9: Artisanal - 20% target, Add KBAs
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in existing KBAs and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_kbas <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_KBAs) %>% # need to add
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_artisanal_kbas <- solve(prob_artisanal_kbas)
## Plot the solution to see what it looks like
plot(sprob_artisanal_kbas,  main = c("Artisanal- 20% Targets, Existing KBAs"))
```




