---
title: "scenario_0_boundary_penalty"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)
library(patchwork)
```


## Boundary Penalty Sensitivity Analysis using Combined Fishing Pressure 
To find an efficient boundary penalty (also called a boundary length modifier in Marxan) to use in the model, we used the combined fishing pressure cost layer and ran the model a number of times. We started with a boundary penalty of 0 and exponentially increased the penalty up to 1 and then plotted the trade-off between total reserve system boundary length and average area of the reserve system for each of these runs to find the most efficient bounadary value. For our model it was 0.000001. 

### **STEP 1: Assign planning units, cost layer, and conservation feature to use in the model runs**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_0 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

# ### plot to make sure it looks okay
plot(pu_0)
```


#### Cost Layer
```{r}
### pull in the combined fishing pressure tif file from costs folder in repo
cost_combined_0 <- raster(here("set_up", "final_rasters", "costs", "combined_stack_sum.tif"))

### plot to make sure it looks okay
plot(cost_combined_0)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack_) <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack_)
```

### Additional Files (Locked-in & Locked-out)
```{r}
# MPAs
### read in existing mpas raster
exist_mpas_0 <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas_0)
```

#### **Step 2: Time to start doing model runs with different boundary penalty values**

The following runs were using combined fishing pressure as a cost with 20% targets and MPA's locked-in.

#### Boundary Penalty: 0
```{r}

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20_0 <- problem(cost_combined_0, features = features_stack_0) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_combined_20_0 <- solve(prob_combined_20_0)
## Plot the solution to see what it looks like
plot(sprob_combined_20_0,  main = c("combined- 0 boundary"))

#Save plot
writeRaster(sprob_combined_20_0, here("final_results", "boundary_penalty", "sprob_combined_20_6hr_blm0.tif"), overwrite = TRUE)
```


#### Boundary Penalty: 0.000003
```{r}

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20_000003 <- problem(cost_combined_0, features = features_stack_0) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000003)

## Solve problem
sprob_combined_20_000003 <- solve(prob_combined_20_000003)
## Plot the solution to see what it looks like
plot(sprob_combined_20_000003,  main = c("combined - 0.000003 boundary"))

#Save plot
writeRaster(sprob_combined_20_01, here("final_results", "boundary_penalty", "sprob_combined_20_6hr_blm0.000003.tif"), overwrite = TRUE)

```


#### Boundary Penalty: 0.000002
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20_000002 <- problem(cost_combined_0, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000002)

## Solve problem
sprob_combined_20_000002 <- solve(prob_combined_20_000002)
## Plot the solution to see what it looks like
plot(sprob_combined_20_000002,  main = c("combined- 0.000002 Boundary"))

#Save plot
writeRaster(sprob_combined_20_000002, here("final_results", "boundary_penalty", "sprob_combined_20_6hr_blm0.000002.tif"), overwrite = TRUE)

```


#### Boundary Penalty: 0.0000007
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20_000007 <- problem(cost_combined_0, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000007)

## Solve problem
sprob_combined_20_000007 <- solve(prob_combined_20_000007)
## Plot the solution to see what it looks like
plot(sprob_combined_20_000007,  main = c("combined- 0.000007 boundary"))

#Save plot
writeRaster(sprob_combined_20_000007, here("final_results", "boundary_penalty", "sprob_combined_20_6hr_blm0.000007.tif"), overwrite = TRUE)

```


#### Boundary Penalty: 0.0001
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_20_0001 <- problem(cost_combined_0, features = features_stack_4) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  #add_locked_in_constraints(exist_mpas_4) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.0001)

## Solve problem
sprob_combined_20_0001 <- solve(prob_combined_20_0001)
## Plot the solution to see what it looks like
plot(sprob_combined_20,  main = c("combined- 20% Targets"))

#Save plot
writeRaster(sprob_combined_20_0001, (here("final_results", "boundary_penalty", "sprob_combined_20_6hr_blm0.0001tif")), overwrite = TRUE)

```

## Convert tifs to shapefiles to extract BLM data
```{r}
#BLM 0
blm0 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.tif"))
blm0_sf <- rasterToPolygons(blm0, fun = function(sprob_combined_20_6hr_blm0){sprob_combined_20_6hr_blm0==1})
plot(blm0_sf)
shapefile(blm0_sf, here("set_up", "blm_sensitivity", "blm0"))

#BLM 0.000000001
blm0.000000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000000001.tif"))
blm0.000000001_sf <- rasterToPolygons(blm0.000000001, fun = function(sprob_combined_20_6hr_blm0.000000001){sprob_combined_20_6hr_blm0.000000001==1})
plot(blm0.000000001_sf)
shapefile(blm0.000000001_sf, here("set_up", "blm_sensitivity", "blm81"))

#BLM 0.00000001
blm0.00000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.00000001.tif"))
blm0.00000001_sf <- rasterToPolygons(blm0.00000001, fun = function(sprob_combined_20_6hr_blm0.00000001){sprob_combined_20_6hr_blm0.00000001==1})
plot(blm0.00000001_sf)
shapefile(blm0.00000001_sf, here("set_up", "blm_sensitivity", "blm71"))

#BLM 0.0000001
blm0.0000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.0000001.tif"))
blm0.0000001_sf <- rasterToPolygons(blm0.0000001, fun = function(sprob_combined_20_6hr_blm0.0000001){sprob_combined_20_6hr_blm0.0000001==1})
plot(blm0.0000001_sf)
shapefile(blm0.0000001_sf, here("set_up", "blm_sensitivity", "blm61"))

#BLM 0.000001
blm0.000001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000001.tif"))
blm0.000001_sf <- rasterToPolygons(blm0.000001, fun = function(sprob_combined_20_6hr_blm0.000001){sprob_combined_20_6hr_blm0.000001==1})
plot(blm0.000001_sf)
shapefile(blm0.000001_sf, here("set_up", "blm_sensitivity", "blm51"))

#BLM 0.00001
blm0.00001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.00001.tif"))
blm0.00001_sf <- rasterToPolygons(blm0.00001, fun = function(sprob_combined_20_6hr_blm0.00001){sprob_combined_20_6hr_blm0.00001==1})
plot(blm0.00001_sf)
shapefile(blm0.00001_sf, here("set_up", "blm_sensitivity", "blm41"))

#BLM 0.0001
blm0.0001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.0001.tif"))
blm0.0001_sf <- rasterToPolygons(blm0.0001, fun = function(sprob_combined_20_6hr_blm0.0001){sprob_combined_20_6hr_blm0.0001==1})
plot(blm0.0001_sf)
shapefile(blm0.0001_sf, here("set_up", "blm_sensitivity", "blm31"))

#BLM 0.001
blm0.001 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.001.tif"))
blm0.001_sf <- rasterToPolygons(blm0.001, fun = function(sprob_combined_20_6hr_blm0.001){sprob_combined_20_6hr_blm0.001==1})
plot(blm0.001_sf)
shapefile(blm0.001_sf, here("set_up", "blm_sensitivity", "blm21"))

#BLM 0.01
blm0.01 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.01.tif"))
blm0.01_sf <- rasterToPolygons(blm0.01, fun = function(sprob_combined_20_6hr_blm0.01){sprob_combined_20_6hr_blm0.01==1})
plot(blm0.01_sf)
shapefile(blm0.01_sf, here("set_up", "blm_sensitivity", "blm11"))

#BLM 0.1
blm0.1 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.1.tif"))
blm0.1_sf <- rasterToPolygons(blm0.1, fun = function(sprob_combined_20_6hr_blm0.1){sprob_combined_20_6hr_blm0.1==1})
plot(blm0.1_sf)
shapefile(blm0.1_sf, here("set_up", "blm_sensitivity", "blm01"))

#BLM 1
blm1 <- raster(here("final_results", "sprob_combined_20_6hr_blm1.tif"))
blm1_sf <- rasterToPolygons(blm1, fun = function(sprob_combined_20_6hr_blm1){sprob_combined_20_6hr_blm1==1})
plot(blm1_sf)
shapefile(blm1_sf, here("set_up", "blm_sensitivity", "blm1"))

#BLM 0.000003
blm0.000003 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000003.tif"))
blm0.000003_sf <- rasterToPolygons(blm0.000003, fun = function(sprob_combined_20_6hr_blm0.000003){sprob_combined_20_6hr_blm0.000003==1})
plot(blm0.000003_sf)
shapefile(blm0.000003_sf, here("set_up", "blm_sensitivity", "blm53"))

#BLM 0.000005
blm0.000005 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000005.tif"))
blm0.000005_sf <- rasterToPolygons(blm0.000005, fun = function(sprob_combined_20_6hr_blm0.000005){sprob_combined_20_6hr_blm0.000005==1})
plot(blm0.000005_sf)
shapefile(blm0.000005_sf, here("set_up", "blm_sensitivity", "blm55"))

#BLM 0.000007
blm0.000007 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000007.tif"))
blm0.000007_sf <- rasterToPolygons(blm0.000007, fun = function(sprob_combined_20_6hr_blm0.000007){sprob_combined_20_6hr_blm0.000007==1})
plot(blm0.000007_sf)
shapefile(blm0.000007_sf, here("set_up", "blm_sensitivity", "blm57"))

#BLM 0.000002
blm0.000002 <- raster(here("final_results", "sprob_combined_20_6hr_blm0.000002.tif"))
blm0.000002_sf <- rasterToPolygons(blm0.000002, fun = function(sprob_combined_20_6hr_blm0.000002){sprob_combined_20_6hr_blm0.000002==1})
plot(blm0.000002_sf)
shapefile(blm0.000002_sf, here("set_up", "blm_sensitivity", "blm52"))

```

## BLM optimization curve plot
```{r}
#read in csv
blm_sensitivity_all <- read_csv(here("set_up", "blm_sensitivity", "blm_sensitivity_values.csv"))

blm_sensitivity_1 <- blm_sensitivity_all %>% 
  filter(round == "1")

# plot it - average area - first round
ggplot(data = blm_sensitivity_1, aes(x = area_avg_km2, y = perimeter_sum_km)) +
  geom_point() +
  geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
  geom_line() +
  labs(x = "Average Area (km2)",
       y = "Boundary Length Sum (km)",
       title = "Boundary Penalty Sensitivity Analysis")

# plot it - average area - with honed in values
ggplot(data = blm_sensitivity_all, aes(x = area_avg_km2, y = perimeter_sum_km)) +
  geom_point() +
  geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
  geom_line() +
  labs(x = "Average Area (km2)",
       y = "Boundary Length Sum (km)",
       title = "Boundary Penalty Sensitivity Analysis")


# plot it - total area
# ggplot(data = blm_sensitivity, aes(x = total_area_km2, y = perimeter_sum_km)) +
#   geom_point() +
#   geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
#   geom_line() +
#   labs(x = "Total Area (km2)",
#        y = "Boundary Length Sum (km)",
#        title = "BLM Sensitivity Analysis - Total Area")
# 
# # plot it - perimeter to area
# ggplot(data = blm_sensitivity, aes(x = perimeter_sum_km, y = total_area_km2)) +
# #ggplot(data = blm_sensitivity, aes(x = total_area_km2 , y = perimeter_sum_km)) +
#   geom_point() +
#   geom_text(aes(label = BLM), hjust = 1, vjust = 1) +
#   geom_line() +
#   labs(x = "Total Area (km2)",
#        y = "Boundary Length Sum (km)",
#        title = "BLM Sensitivity Analysis - Total Area")

```





