---
title: "Scenario 2: Artisanal Fishing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)

```

## Scenario 2 - Artisanal Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on artisanal fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### STEP 1: Assign planning units, cost layer, and conservation feature

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_2 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))
# 
# ### plot to make sure it looks okay
plot(pu_2)

```


#### Cost Layer
Other scenarios will require more prep to create cost layer, so they will be found in separate Rmarkdowns. Since this scenario is so short, we created it in this Rmarkdown.
```{r}
### pull in the artisanal fishing pressure tif file from costs folder in repo
art_fish_rast <- raster(here("set_up", "final_rasters", "costs", "kernel_d_fishing_moz_v2.tif"))

## now we need to grab the Moz raster we previously created to make sure the artisanal fishing raster is in the right projections
art_fish <- raster::projectRaster(art_fish_rast, pu_2, res = 10000, method = 'ngb')

## create a data frame to see what value ranges we are working with
art_fish_df <- as.data.frame(art_fish)

## plot to see what it looks like
plot(art_fish)
 
## Priortizr does not like cost values of NA or 0 so changed both to 0.1
art_fish[art_fish <= 0] <- 0.1
art_test <- as.data.frame(art_fish)

#art_fish_rast <- subs(pu_2, art_fish_test, by = 'id', which = 'fish_press')

plot(art_fish)

## Since we changed all NAs values to 0.1 we need to mask this to MZ eez since we are now picking up terrestrial areas 
cost_art_fish <- mask(art_fish, mz_rast_id)
plot(cost_art_fish)

## Looks good lets save as a tif
writeRaster(cost_art_fish, here('set_up/final_rasters/cost_art_fish.tif'), overwrite = TRUE)

```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other scenarios (ie scenario_1_area)

#species_stack

#habitat_stack

##combine into 
#features_stack
```

### Additional Files (Locked-in & Locked-out)
```{r}
# List of additional files that will have been loaded into the global environment in other scenarios (ie scenario_1_area)

#exist_mpas

#oil_rast

```

## Let's Run Some Artisanal Fishing Scenarios!

## SCENARIO 1: Baseline - 20% target, Boundary Penalty = 0,  
```{r}
### OBJECTIVE: Minimize artisanal fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal <- problem(cost_art_fish, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 10)

## Solve problem
sprob_artisanal <- solve(prob_artisanal)
## Plot the solution to see what it looks like
plot(sprob_artisanal,  main = c("Artisanal- 20% Targets"))


```



