---
title: "Scenario 3 - Industrial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
#library(gurobi)

```

## Scenario 2 - Industrial Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on indsutrial fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_3 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))
# 
# ### plot to make sure it looks okay
plot(pu_3)
```


#### Cost Layer
```{r}
### pull in the industrial fishing pressure tif file from costs folder in repo
cost_industrial <- raster(here("set_up", "final_rasters", "costs", "cost_industrial_sum.tif"))

### plot to make sure it looks okay
plot(cost_industrial)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas)
```

## **STEP 2: Run Some Industrial Fishing Scenarios!**

**Changing the targets!**

#### SCENARIO 1: Industrial - 20% target, Boundary Penalty = 0,  
```{r}
### OBJECTIVE: Minimize industrial fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_20 <- problem(cost_industrial, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_industrial_20 <- solve(prob_industrial_20)
## Plot the solution to see what it looks like
plot(sprob_industrial_20,  main = c("industrial- 20% Targets"))
```

#### SCENARIO 2: Industrial - 10% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 10% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_industrial_10 <- problem(cost_industrial, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_industrial_10 <- solve(prob_industrial_10)
## Plot the solution to see what it looks like
plot(sprob_industrial_10,  main = c("Industrial- 10% Targets"))
```


#### SCENARIO 3: Industrial - 50% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 50% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_industrial_50 <- problem(cost_industrial, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_industrial_50 <- solve(prob_industrial_50)
## Plot the solution to see what it looks like
plot(sprob_industrial_50,  main = c("Industrial- 50% Targets"))
```


