---
title: "Scenario 1: Area Model Run"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)

```

## Scenario 1 - Baseline
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats using a uniform cost for all planning units. For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
### reading in the baseline MZ EEZ raster (with cell ids) 
pu <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

### plot to make sure it looks okay
plot(pu)
```

#### Cost Layer
```{r}
### pull in the cost layer created in setup step 9
cost_area <- raster(here("set_up", "final_rasters", "costs", "area_cost.tif"))

### plot to make sure it looks okay
plot(cost_area)
```


#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
### pull in species stack that we created in set-up step 3
features_stack <- brick(here("set_up", "final_rasters", "conservation_features","all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas)
```
### **STEP 2: Run Models and Sensitivity Analysis**

#### RUN 1: Baseline - 20% target, Boundary Penalty = 0,  
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_area_20 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_area_20 <- solve(prob_area_20)
## Plot the solution to see what it looks like
plot(sprob_area_20,  main = c("Area- 20% Targets"))

#Write Raster
writeRaster(sprob_area_20, here("final_results", "scenario_1_area",  "sprob_area_20_6hr_blm0.tif"), overwrite = TRUE)
```

Final plot - reading in
```{r}
## read in scenario if already saved (so don't have to run model again)
area_20_6hr_blm0 <- raster(here("final_results", "scenario_1_area", "sprob_area_20_6hr_blm0.tif"))
plot(area_20_6hr_blm0)

area_20_6hr_blm0_df <- area_20

ggplot()+
  geom_raster(data = area_20_6hr_blm0)
```



#### RUN 2: Baseline - 10% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 10% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_area_10 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_area_10 <- solve(prob_area_10)
## Plot the solution to see what it looks like
plot(sprob_area_10,  main = c("Area- 10% Targets"))

#Write Raster
writeRaster(sprob_area_10, here("final_results", "scenario_1_area",  "sprob_area_10_6hr_blm0.tif"), overwrite = TRUE)
```


#### RUN 3: Baseline - 50% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 50% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_area_50 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_area_50 <- solve(prob_area_50)
## Plot the solution to see what it looks like
plot(sprob_area_50,  main = c("Area- 50% Targets"))

#Write Raster
writeRaster(sprob_area_50, here("final_results", "scenario_1_area",  "sprob_area_50_6hr_blm0.tif"), overwrite = TRUE)

```


#### SCENARIO 1: Baseline - 20% target, Boundary Penalty = 0.000001,  
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_area_20 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_area_50 <- solve(prob_area_50)
## Plot the solution to see what it looks like
plot(sprob_area_50,  main = c("Area- 50% Targets"))

#Write Raster
writeRaster(sprob_area_50, here("final_results", "scenario_1_area",  "sprob_area_50_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### Baseline -30% target, Boundary Penaly = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 30% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_area_30 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_area_30 <- solve(prob_area_30)
## Plot the solution to see what it looks like
plot(sprob_area_30,  main = c("Area- 30% Targets"))

#Write Raster
writeRaster(sprob_area_30, here("final_results", "scenario_1_area",  "sprob_area_30_6hr_blm0.tif"), overwrite = TRUE)
```



```{r}
# Sensitivity analysis with MPA's not locked in
prob_area_20_nompa <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_area_20_nompa <- solve(prob_area_20_nompa)
## Plot the solution to see what it looks like
plot(sprob_area_20_nompa,  main = c("Area- 20% Targets without MPA's locked in"))

#Write Raster
writeRaster(sprob_area_20_nompa, here("final_results", "scenario_1_area",  "sprob_area_20_6hr_blm0_nompa.tif"), overwrite = TRUE)
```



