---
title: "Scenario 5 - Inverse Combined Fishing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)
```

## Scenario 5 - Inverse Combined Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while targeting the areas with high fishing pressure  For this run, we used the inverse of the combined fishing pressure layer as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_5 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

# ### plot to make sure it looks okay
plot(pu_5)
```

#### Cost Layer
```{r}
### pull in the combined fishing pressure tif file from costs folder in repo
cost_combined_inverse <- raster(here("set_up", "final_rasters", "costs", "combined_inverse_simple.tif"))

### plot to make sure it looks okay
plot(cost_combined_inverse)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack_5 <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack_5)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas_5 <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas_5)
```

## **STEP 2: Run Some inverse combined Fishing Scenarios!**

**Changing the targets!**

#### SCENARIO 1: combined - 20% target, Boundary Penalty = 0,  
```{r}
### OBJECTIVE: Minimize combined fishing conflict while meeting 20% representation targets, locking in MPAs, and a boundary penalty of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_combined_inverse_10 <- problem(cost_combined_inverse, features = features_stack_5) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas_5) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_combined_inverse_10 <- solve(prob_combined_inverse_10)
## Plot the solution to see what it looks like
plot(sprob_combined_inverse_10,  main = c("inverse- 10% Targets"))

#Write Raster
writeRaster(sprob_combined_inverse_10, here("final_results", "scenario_5_inverse",  "sprob_combined_inverse_10_6hr_blm0.000001.tif"), overwrite = TRUE)

```