---
title: "Scenario 1: Area Model Run"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)

```

## Scenario 1 - Baseline
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats using a uniform cost for all planning units. For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### STEP 1: Assign planning units, cost layer, and conservation feature

#### Planning Unit
```{r}
### reading in the baseline MZ EEZ raster (with cell ids) 
pu <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

### plot to make sure it looks okay
plot(pu)

```

#### Cost Layer
Other scenarios will require more prep to create cost layer, so they will be found in separate Rmarkdowns. Since this scenario is so short, we created it in this Rmarkdown. 
```{r}
### pull in the initial MZ EEZ raster to start with
area_raster <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))

### assign a uniform cost value of 0.5 to all cells since the area of the planning units are the same. 
### We assigned a value of 0.5 since we will transform cost layers in future simulations to be between 0-1 and this 0.5 is a middle value in this range 
values(area_raster) <- 0.5

### mask the area raster to just the MZ EEZ, so all cells outside the EEZ are assigned NA AND rename it to cost_area
cost_area <- mask(area_raster, pu)

### check it by plotting!
plot(cost_area)

### write raster
writeRaster(cost_area, here('set_up/final_rasters/area_cost.tif'), overwrite = TRUE)

```


#### Conservation Features
This includes all species distributions and habitat rasters

**Creating Species Stack**
```{r}
### pull in species stack that we created in set-up step 3

species_stack <- brick(here("set_up", "final_rasters", "conservation_features", "species_stack.tif"))

plot(species_stack)

```



Creating Final Conservation Feature 
```{r}

```


### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

```


## SCENARIO 1: Baseline - 20% target, Boundary Penalty = 0,  
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_area <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_area <- solve(prob_area)
## Plot the solution to see what it looks like
plot(sprob_area,  main = c("Area- 20% Targets"))


```


## SCENARIO 2: Baseline - 10% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_area_10 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_area_10 <- solve(prob_area_10)
## Plot the solution to see what it looks like
plot(sprob_area_10,  main = c("Area- 10% Targets"))
```

## SCENARIO 3: Baseline - 30% target, Boundary Penalty = 0 
```{r}
### OBJECTIVE: Minimize area while meeting 20% representation targets, locking in MPAs, locking out oil rigs, and a boundary penatly of 0

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_area_30 <- problem(cost_area, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_area_30 <- solve(prob_area_30)
## Plot the solution to see what it looks like
plot(sprob_area_30,  main = c("Area- 10% Targets"))
```
















































