---
title: "Scenario 3 - Industrial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)
```

## Scenario 3 - Industrial Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on indsutrial fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_3 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))
# 
# ### plot to make sure it looks okay
plot(pu_3)
```


#### Cost Layer
```{r}
### pull in the industrial fishing pressure tif file from costs folder in repo
cost_industrial <- raster(here("set_up", "final_rasters", "costs", "cost_industrial_sum.tif"))

### plot to make sure it looks okay
plot(cost_industrial)
```

#### Conservation Features
This includes all species distributions and habitat rasters

```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack_3 <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack_3)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas_3 <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas_3)
```

## **STEP 2: Run Some Industrial Fishing Scenarios with no boundary penalty**

**Changing the targets!**

#### RUN 1: Industrial - 10% target, Boundary Penalty = 0 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_industrial_10 <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_industrial_10 <- solve(prob_industrial_10)


## Plot the solution to preview
plot(sprob_industrial_10,  main = c("Industrial- 10% Targets"))

#Save plot
writeRaster(sprob_industrial_10, here("final_results", "scenario_3_industrial", "sprob_industrial_10_6hr_blm0.tif"), overwrite = TRUE)
```

#### RUN 2: Industrial - 20% target, Boundary Penalty = 0,  
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_20 <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) 

## Solve problem
sprob_industrial_20 <- solve(prob_industrial_20)


## Plot the solution to preview
plot(sprob_industrial_20,  main = c("industrial- 20% Targets"))

#@ Save plot
writeRaster(sprob_industrial_20, here("final_results", "scenario_3_industrial", "sprob_industrial_20_6hr_blm0.tif"), overwrite = TRUE)

```


#### RUN 3: Industrial - 30% Target, Boundary Penalty = 0
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_30 <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 21600)

## Solve problem
sprob_industrial_30 <- solve(prob_industrial_30)
## Plot the solution to see what it looks like
plot(sprob_industrial_30,  main = c("Industrial- 30% Targets"))

#Save plot
writeRaster(sprob_industrial_30, here("final_results", "scenario_3_industrial", "sprob_industrial_30_6hr_blm0.tif"), overwrite = TRUE)
```


#### RUN 4: Industrial - 50% target, Boundary Penalty = 0 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_industrial_50 <- problem(cost_industrial, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120)

## Solve problem
sprob_industrial_50 <- solve(prob_industrial_50)

## Plot the solution to preview
plot(sprob_industrial_50,  main = c("Industrial- 50% Targets"))

## Save plot
writeRaster(sprob_industrial_30, here("final_results", "scenario_3_industrial", "sprob_industrial_50_6hr_blm0.tif"), overwrite = TRUE)
```

---


## **Run Some Industrial Fishing Scenarios with efficient boundary penalty (1x10^-6)**

#### Run 5: Industrial - 10% target, Boundary Penaly = 0.000001
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_10_blm <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_industrial_10_blm <- solve(prob_industrial_10_blm)

## Plot the solutionto preview
plot(sprob_industrial_10_blm,  main = c("industrial- 10% targets with boundary penalty"))

## Save plot
writeRaster(sprob_industrial_10_blm, here("final_results", "scenario_3_industrial", "sprob_industrial_10_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### RUN 6: Industrial - 20% target, Boundary Penalty = 0.000001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_20_blm <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_industrial_20_blm <- solve(prob_industrial_20_blm)

## Plot the solution to preview
plot(sprob_industrial_20_blm,  main = c("industrial- 20% targets with boundary penalty"))

## Save plot
writeRaster(sprob_industrial_20_blm, here("final_results", "scenario_3_industrial", "sprob_industrial_20_6hr_blm0.000001.tif"), overwrite = TRUE)
```

#### RUN 7: Industrial - 30% target, Boundary Penalty = 0.000001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_30_blm <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_industrial_30_blm <- solve(prob_industrial_30_blm)

## Plot the solution to preview
plot(sprob_industrial_30_blm,  main = c("industrial- 30% Targets with boundary penalty"))

## Save plot
writeRaster(sprob_industrial_30_blm, here("final_results", "scenario_3_industrial", "sprob_industrial_30_6hr_blm0.000001.tif"), overwrite = TRUE)
```

#### RUN 8: Industrial - 50% target, Boundary Penalty = 0.000001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_50_blm <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_industrial_50_blm <- solve(prob_industrial_50_blm)

## Plot the solution to preview
plot(sprob_industrial_50_blm,  main = c("industrial- 50% targets with boundary penalty"))

## Save plot
writeRaster(sprob_industrial_50_blm, here("final_results", "scenario_3_industrial", "sprob_industrial_50_6hr_blm0.000001.tif"), overwrite = TRUE)
```

**Now, time to run some sensitivity analysis with MPAs**

#### Run 9: Industrial - 20% targets, boundary = 0 and no MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_20_nompa <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) 

## Solve problem
sprob_industrial_20_nompa <- solve(prob_industrial_20_nompa)


## Plot the solution to preview
plot(sprob_industrial_20_nompa,  main = c("industrial- 20% targets with no MPA"))

## Save plot
writeRaster(sprob_industrial_20_nompa, here("final_results", "sensitivity", "mpas", "sprob_industrial_20_6hr_blm0_nompa.tif"), overwrite = TRUE)

```


#### Run 10: Industrial - 20% targets, boundary = 0.000001 and no MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_industrial_20_blm_nompa <- problem(cost_industrial, features = features_stack_3) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas_3) %>%
  add_gurobi_solver(gap = 0.01, time_limit = 120) %>% 
  add_boundary_penalties(0.000001)

## Solve problem
sprob_industrial_20_blm_nompa <- solve(prob_industrial_20_blm_nompa)


## Plot the solution to preview
plot(sprob_industrial_20_blm_nompa,  main = c("industrial- 20% targets with boundary and no MPA"))

## Save plot
writeRaster(sprob_industrial_20_nompa, here("final_results", "sensitivity", "mpas", "sprob_industrial_20_6hr_blm0.000001_nompa.tif"), overwrite = TRUE)

```



