---
title: "Scenario 2: Artisanal Fishing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(prioritizr)
library(rgdal)
library(gurobi)
```

## Scenario 2 - Artisanal Fishing Pressure 
The objective of this scenario was to protect important areas for sharks and rays based on their distribution from IUCN and location of suitable habitats while minimizing the impact on artisanal fisherman.  For this run, we used area as the cost, set targets of 20% for all conservation features, locked in planning units that included existing MPAs, and set a boundary penalty to zero. 

Step 1 reads in all the necessary data
Step 2 run to see results of varying targets from 10-50%
Step 3 to see varying targets using a boundary penalty
Step 4 to run sensitivity analysis of locking in vs locking out MPAs

----

### **STEP 1: Assign planning units, cost layer, and conservation feature**

#### Planning Unit
```{r}
# ### reading in the baseline MZ EEZ raster (with cell ids) 
pu_2 <- raster(here("set_up", "final_rasters", "mz_eez_templates", "mz_rast_id.tif"))
# 
# ### plot to make sure it looks okay
plot(pu_2)
```


#### Cost Layer
```{r}
### pull in the artisanal fishing pressure tif file from costs folder in repo
cost_artisanal <- raster(here("set_up", "final_rasters", "costs", "cost_art_fish.tif"))

### plot to make sure it looks okay
plot(cost_artisanal)
```

#### Conservation Features
This includes all species distributions and habitat rasters
```{r}
# List of conservation features that will have been loaded into the global environment in other
### pull in species stack that we created in set-up step 3
features_stack <- brick(here("set_up", "final_rasters", "conservation_features", "all_features_stack.tif"))

### plot to make sure it looks okay
plot(features_stack)
```

### Additional Files (Locked-in & Locked-out)
```{r}
### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas)
```

### Set the some of the global model parameters like gap, time limit, and efficient boundary penalty to use in soem runs:
```{r}
# if you decide to change the gap, time-limit or boundary penalty value, change it here and rerun the subsequent model runs

gap <- 0.01
time_limit <- 21600
bp <- 0.000001
```


### **STEP 2: Run Models with varying targets**

#### RUN 1: Artisanal - 10% target, Boundary Penalty = 0 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_10 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_artisanal_10 <- solve(prob_artisanal_10)

## Plot the solution to see what it looks like
plot(sprob_artisanal_10,  main = c("Artisanal- 10% Targets"))

## Save as tif
writeRaster(sprob_artisanal_10, here("final_results", "scenario_2_artisanal", "sprob_artisanal_10_6hr_blm0.tif"), overwrite = TRUE)
```


#### RUN 2: Artisanal - 20% target, Boundary Penalty = 0
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal_20 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_artisanal_20 <- solve(prob_artisanal_20)

## Plot the solution to see what it looks like
plot(sprob_artisanal_20,  main = c("Artisanal- 20% Targets"))

## Save as tif
writeRaster(sprob_artisanal_20, here("final_results", "scenario_2_artisanal", "sprob_artisanal_20_6hr_blm0.tif"), overwrite = TRUE)
```


#### RUN 3: Artisanal - 30% target, Boundary Penalty = 0
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal_30 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_artisanal_30 <- solve(prob_artisanal_30)

## Plot the solution to see what it looks like
plot(sprob_artisanal_30,  main = c("Artisanal- 30% Targets"))

## Save as tif
writeRaster(sprob_artisanal_30, here("final_results", "scenario_2_artisanal", "sprob_artisanal_30_6hr_blm0.tif"), overwrite = TRUE)
```


#### RUN 4: Artisanal - 50% target, Boundary Penalty = 0 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal_50 <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_artisanal_50 <- solve(prob_artisanal_50)

## Plot the solution to see what it looks like
plot(sprob_artisanal_50,  main = c("Artisanal- 50% Targets"))

#Save plot
writeRaster(sprob_artisanal_50, here("final_results", "scenario_2_artisanal", "sprob_artisanal_50_6hr_blm0.tif"), overwrite = TRUE)
```


### **STEP 3: Run Models with efficient boundary penalty (0.000001)**

#### RUN 5: Artisanal - 10% target, Boundary Penalty = 0.000001
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_10_blm <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.1) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(penalty = bp)

## Solve problem
sprob_artisanal_10_blm <- solve(prob_artisanal_10_blm)

## Plot the solution to see what it looks like
plot(sprob_artisanal_10_blm,  main = c("Artisanal- 10% Targets (with boundary penalty)"))

## Save as tif
writeRaster(sprob_artisanal_10_blm, here("final_results", "scenario_2_artisanal", "sprob_artisanal_10_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### RUN 6: Artisanal - 20% target, Boundary Penalty = 0.00001 
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objective
prob_artisanal_20_blm <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(penalty=bp)

## Solve problem
sprob_artisanal_20_blm  <- solve(prob_artisanal_20_blm)

## Plot the solution to see what it looks like
plot(sprob_artisanal_20_blm,  main = c("Artisanal- 20% Targets (with boundary penalty)"))

## Save as tif
writeRaster(sprob_artisanal_20_blm, here("final_results", "scenario_2_artisanal", "sprob_artisanal_20_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### RUN 7: Artisanal - 30% target, Boundary Penalty = 0.000001
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive

prob_artisanal_30_blm <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.3) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(penalty=bp)

## Solve problem
sprob_artisanal_30_blm  <- solve(prob_artisanal_30_blm )

## Plot the solution to see what it looks like
plot(sprob_artisanal_30_blm ,  main = c("Artisanal- 30% Targets (with boundary penalty)"))

## Save as tif
writeRaster(sprob_artisanal_30_blm, here("final_results", "scenario_2_artisanal", "sprob_artisanal_30_6hr_blm0.000001.tif"), overwrite = TRUE)
```


#### RUN 8: Artisanal - 50% target, Boundary Penalty = 0.000001
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_50_blm <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.5) %>%
  add_locked_in_constraints(exist_mpas) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(penalty=bp)

## Solve problem
sprob_artisanal_50_blm  <- solve(prob_artisanal_50_blm )

## Plot the solution to see what it looks like
plot(sprob_artisanal_50_blm ,  main = c("Artisanal- 50% Targets (with boundary penalty)"))

## Save as tif
writeRaster(sprob_artisanal_50_blm, here("final_results", "scenario_2_artisanal", "sprob_artisanal_50_6hr_blm0.000001.tif"), overwrite = TRUE)
```


### **STEP 4: Run Models with MPA sensitivity analysis**

#### RUN 9: Artisanal - 20% targets, boundary penalty = 0, and no MPAs locked in
```{r}

## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_20_nompa <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit)

## Solve problem
sprob_artisanal_20_nompa  <- solve(prob_artisanal_20_nompa )

## Plot the solution to see what it looks like
plot(sprob_artisanal_20_nompa ,  main = c("Artisanal - 20% Targets without MPA's locked in"))

## Save as tiff
writeRaster(sprob_artisanal_20_nompa, here("final_results", "scenario_2_artisanal",  "sprob_artisanal_20_6hr_blm0_nompa.tif"), overwrite = TRUE)
```


#### RUN 10: Artisanal - 20% targets, boundary penalty = 0.000001, and no MPAs locked in
```{r}
## Define the problem with the cost, conservation features, targets, locked in and locked out constraints using the minimum set objecive
prob_artisanal_20_blm_nompa <- problem(cost_artisanal, features = features_stack) %>%
  add_min_set_objective() %>%
  add_relative_targets(0.2) %>%
  add_gurobi_solver(gap = gap, time_limit = time_limit) %>% 
  add_boundary_penalties(bp)

## Solve problem
sprob_artisanal_20_blm_nompa  <- solve(prob_artisanal_20_blm_nompa )

## Plot the solution to see what it looks like
plot(sprob_artisanal_20_blm_nompa ,  main = c("Artisanal - 20% Targets without MPA's locked in"))

## Save as tiff
writeRaster(sprob_artisanal_20_blm_nompa, here("final_results", "scenario_2_artisanal",  "sprob_artisanal_20_6hr_blm0.000001_nompa.tif"), overwrite = TRUE)
```






