---
title: "Bruv Validation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(fasterize)
```

#### **STEP 1:Read in MZ EZZ files and species distribution for Grey Reef Shark (Carcharhinus amblyrhynchos)**
```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))
plot(mz_rast_id)

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## read in grey reef shark iucn species distribution made in 2_iucn_species.Rmd
iucn_grey_reef_shark <- raster(here("set_up/final_rasters/conservation_features/iucn_species/species_Carcharhinus amblyrhynchos.tif"))
plot(iucn_grey_reef_shark)
```

#### **STEP 2: Next, run `5_bruv.Rmd` file to get the bruv files in your environment**

#### **STEP 3: Read in priorization results for scenario 4 and 5**
```{r}
combined_path <- here("blm_sensitivity")
combined_files <- list.files(combined_path, full.names = TRUE)

combined_20_blm0.000001 <- raster(combined_files[4])
plot(combined_20_blm0.000001)
combined_20_blm0 <- raster(combined_files[14])
plot(combined_20_blm0)

#combined_20_blm0.000001[combined_20_blm0.000001 == 0] <- NA

inverse_path <- here("final_results", "scenario_5_inverse")
inverse_files <- list.files(inverse_path, full.names = TRUE)

inverse_20_blm0.000001 <- raster(inverse_files[3])
plot(inverse_20_blm0.000001)
inverse_20_blm0 <- raster(inverse_files[4])
plot(inverse_20_blm0)

### read in existing mpas raster
exist_mpas <- raster(here("set_up", "final_rasters", "planning_unit", "exist_mpa.tif"))

### plot to make sure it looks okay
plot(exist_mpas)

```

#### **STEP 4: Filter the focal species in the BRUV data**
```{r}
grey_reef_shark_df <- wcs_bruv_sf_maxn %>% 
  filter(genus_species == "Carcharhinus amblyrhynchos")
  
```

#### **STEP 5: Create a function to create interactive map with final priorization model, existing mpas, BRUV data, and IUCN distribution**
```{r}
create_bruv_map <- function(final_raster, mpas, bruv, iucn_distribution){

  final_raster[final_raster == 0] <- NA
  
  tmap_mode("view") # Set to interactive viewing

  tm_shape(iucn_grey_reef_shark)+
    tm_raster("species_Carcharhinus_amblyrhynchos", palette="darkslategray3", alpha = 0.8,
              legend.show = F)+
  tm_shape(final_raster)+
    tm_raster(final_raster[1], palette="navy", alpha = 0.65,legend.show = F)+
  tm_shape(mz_eez_sf)+
    tm_borders("black", lwd = .5) +
  tm_shape(exist_mpas)+
    tm_raster(exist_mpas[1], palette = "lightcoral", legend.show = F)+
  tm_shape(grey_reef_shark_df) +
    tm_dots("genus_species", clustering = T,  legend.show = FALSE, size = 0.5)+
  tm_add_legend("fill",
              col = c("navy", "lightcoral", "darkslategray3"),
	            border.col = "black",
	            labels = c('Suggested Areas','Existing MPAs', 'IUCN Distribution'),
	            title="Prioritization Model & BRUV data")
}

create_bruv_map(combined_20_blm0.000001, exist_mpas, grey_reef_shark_df, iucn_grey_reef_shark)
create_bruv_map(combined_20_blm0, exist_mpas, grey_reef_shark_df, iucn_grey_reef_shark)

create_bruv_map(inverse_20_blm0.000001, exist_mpas, grey_reef_shark_df, iucn_grey_reef_shark)
create_bruv_map(inverse_20_blm0, exist_mpas, grey_reef_shark_df, iucn_grey_reef_shark)
```


```{r}
create_all_bruv_map <- function(final_raster, mpas, bruv1, bruv2, bruv3){

  final_raster[final_raster == 0] <- NA
  
  tmap_mode("view") # Set to interactive viewing

  tm_shape(final_raster)+
    tm_raster(final_raster[1], palette="navy",alpha = 0.65, legend.show = F)+
  tm_shape(mz_eez_sf)+
    tm_borders("black", lwd = .5) +
  tm_shape(exist_mpas)+
    tm_raster(exist_mpas[1], palette = "lightcoral", legend.show = F)+
  tm_shape(bruv1) +
    tm_dots("genus_species", clustering = T,  legend.show = FALSE, size = 0.5)+
  tm_shape(bruv2) +
    tm_dots("sample_code", clustering = T,  legend.show = FALSE, size = 0.5)+
  tm_shape(bruv3) +
    tm_dots("sample_code", clustering = T,  legend.show = FALSE, size = 0.5)+
    tmap_options(max.categories = 220)+
    tm_add_legend("fill",
              col = c("navy", "lightcoral"),
	            border.col = "black",
	            labels = c('Suggested Areas','Existing MPAs'),
	            title="Prioritization Model & BRUV data")
}

create_all_bruv_map(combined_20_blm0.000001, exist_mpas, wcs_bruv_sf_maxn, popmr_bruv_sf,nm_bruv_sf)
create_all_bruv_map(combined_20_blm0, exist_mpas, wcs_bruv_sf_maxn, popmr_bruv_sf,nm_bruv_sf)

create_all_bruv_map(inverse_20_blm0.000001, exist_mpas, wcs_bruv_sf_maxn, popmr_bruv_sf,nm_bruv_sf)
create_all_bruv_map(inverse_20_blm0, exist_mpas, wcs_bruv_sf_maxn, popmr_bruv_sf,nm_bruv_sf)

```


























