---
title: "10_inverse_combined_fishing_layer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(spatialEco)
library(tmap)
```

# Cost: Inverse Fishing Pressure
In order to maximize the protection of sharks and rays from areas of high fishing pressure we also created an inverse of this combined fishing pressure cost layer. For this, we used the inverse of the combined fishing cost layer so that our model prioritized protection of areas with the highest fishing pressure and rescaled it between 0 and 1, where 0 was areas with the lowest fishing pressure and 1 was areas with the highest fishing pressure.

#### **STEP 1: Read in data**
```{r}
## read in combined fishing pressure 
combined_inverse_simple <- raster(here("set_up", "final_rasters", "costs", "combined_stack_sum.tif"))
plot(combined_inverse_simple)

## read in summed industrial fishing pressure (not rescaled)
industrial_cumulative <- brick(here('set_up/final_rasters/costs/cost_industrial_sum_unscaled.tif'))
plot(industrial_cumulative)

## read in artisanal fishing pressure
artisanal <- raster(here('set_up/final_rasters/costs/cost_art_fish.tif'))
plot(artisanal)
```

#### **STEP 2: Created Inverse Layer**

Method 1 - Simple; just inverse the other combined layer
```{r}
## invert raster layer using `raster.invert`
combined_inverse_simple_rast <- raster.invert(combined_inverse_simple)

## check it out!
plot(combined_inverse_simple_rast)

## save it
writeRaster(combined_inverse_simple_rast, here("set_up", "final_rasters", "costs", "combined_inverse_simple.tif"), overwrite = TRUE)

```

Method 2 - Cumulative; use the unrescaled industrial and combine with artisanal (not used in our analysis)
```{r}
## create a raster stack of industrial and artisanal
combined_inverse_cumulative <- stack(industrial_cumulative, artisanal)

## plot it
plot(combined_inverse_cumulative)

## add layers together
combined_sum_2 <- calc(combined_inverse_cumulative, fun = sum)

## invert summed raster layer
combined_sum_inverse <- raster.invert(combined_sum_2)

## plot it!
plot(combined_sum_inverse)

## save it
writeRaster(combined_sum_inverse, here("set_up", "final_rasters", "costs", "combined_inverse_complex.tif"), overwrite = TRUE)
```

Wow another interactive map!
```{r}
tmap_mode("view") # Set to interactive viewing

tm_shape(combined_inverse_simple_rast) +
  tm_basemap("Stamen.Terrain")+
  tm_raster("combined_stack_sum", alpha=0.9, palette = "viridis")
  
```

Outputs:

 - `combined_inverse_simple.tif` = inverse raster of combined fishing pressure (simple method)
 - `combined_inverse_complex.tif` = inverse raster of combined fishing pressure (complex method)

End script.
