---
title: "Distance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(fasterize)
```

## Upper Shelf Raster
Depth at 50m (hopefully)

#### **STEP 1: Read in Mozambique EEZ and Shapefile**
```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## Add a buffer of 10 km to make sure cells on the outer edges of the EEZ are included
mz_eez_buffer_sf <- sf::st_buffer(mz_eez_sf, dist = 10000)
```

#### **STEP 2: Read in Bathymetry Data (rachel edited?)**
```{r}
## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
bathy_path <- 'G:/group_project/data/habitats/bathy_classes/rachel_edited'
bathy_shape <- list.files(bathy_path, full.names = TRUE) 

bathy_sf <- sf::read_sf(bathy_shape[8])

bathy_df <- bathy_sf %>%
  as.data.frame() %>% 
  select(-geometry)

## check coordinate system
st_crs(bathy_sf) == st_crs(mz_eez_sf)

### create dataframe that only contains species and changes to correct crs
bathy_sf_reproj <- bathy_sf %>% 
  st_transform(., crs = st_crs(mz_eez_sf)) 

### check crs to make sure it matches
st_crs(bathy_sf_reproj) == st_crs(mz_eez_sf)
```

#### **STEP 3: Create a raster from the sf files**
```{r}
## check the class of sf
class(bathy_sf_reproj)

## change to a sp object
bathy_sp <- as(bathy_sf_reproj, 'Spatial')

## check the class to make sure it worked
class(bathy_sp)

## double check that this matches your RASTER file (which should be the same as the sf)
crs(bathy_sp)

## make a raster 
bathy_rast <- raster::rasterize(bathy_sp, mz_rast_id, field = "gridcode")

## plot to see what it looks like
plot(bathy_rast)
```

#### **STEP 4: Reclassify the depth classes so only show the Upper Shelf (~50m depth)**
Not sure if this is correct, but there is no number for depth just classifications -- change if needed
```{r}
## create a matrix of needed classes
m <- c(0.00, 7, NA, # made NA (instead of 0) for distance decay 
       8, 8, 1)

### wrap the vector into a matrix by rows
rcl_mtx <- matrix(m, ncol=3, byrow=TRUE)

## reclassify with the new matrix
bathy_reclass <- reclassify(bathy_rast, rcl_mtx)

## plot it to check
plot(bathy_reclass)

### save it
writeRaster(bathy_reclass, here('set_up/final_rasters/costs/bathy_rast.tif'), overwrite = TRUE)
```


## Distance Decay with 50m priorization (from above)

```{r}
## use the raster::distance() function, which calculates a distance for every NA cell to the nearest non-NA cell.
dist_rast <- raster::distance(bathy_reclass) %>% 
  mask(mz_eez_sf) # mask so the land distances are not calculated

## plot it!
plot(dist_rast)

## created a dataframe to check it out
dist_df <- dist_rast %>%
  rasterToPoints() %>%
  as.data.frame()

### save it
writeRaster(dist_rast, here('set_up/final_rasters/costs/dist_rast.tif'), overwrite = TRUE)

```






