---
title: "Distance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(fasterize)
```

# Cost: Distance from shore
In addition to the five scenarios, we also ran the model to prioritize areas closer to shore, by using a distance decay from 40m depth and beyond as a cost. This was to factor the management of MPAs into the analysis. Management of MPAs by CCPs and artisanal fishers will require that the MPAs are within the depth range that artisanal fishers typically travel (0 to 50m depth). If MPAs are managed by government officials, MPAs closer to shore will be less costly to manage as they will require less time and fuel to reach for patrolling.

#### **STEP 1: Read in Mozambique EEZ and Shapefile**
```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## Add a buffer of 10 km to make sure cells on the outer edges of the EEZ are included
mz_eez_buffer_sf <- sf::st_buffer(mz_eez_sf, dist = 10000)
```

#### **STEP 2: Read in Bathymetry Data**
```{r}
## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
bathy_path <- 'G:/group_project/data/habitats/bathy_classes/raw_data'
bathy_shape <- list.files(bathy_path, full.names = TRUE) 

bathy_sf <- sf::read_sf(bathy_shape[8])

bathy_df <- bathy_sf %>%
  as.data.frame() %>% 
  select(-geometry)

## check coordinate system
st_crs(bathy_sf) == st_crs(mz_eez_sf)

### create dataframe that only contains species and changes to correct crs
bathy_sf_reproj <- bathy_sf %>% 
  st_transform(., crs = st_crs(mz_eez_sf)) 

### check crs to make sure it matches
st_crs(bathy_sf_reproj) == st_crs(mz_eez_sf)
```

#### **STEP 3: Create a raster from the sf files**
```{r}
## check the class of sf
class(bathy_sf_reproj)

## change to a sp object
bathy_sp <- as(bathy_sf_reproj, 'Spatial')

## check the class to make sure it worked
class(bathy_sp)

## double check that this matches your RASTER file (which should be the same as the sf)
crs(bathy_sp)

## make a raster 
bathy_rast <- raster::rasterize(bathy_sp, mz_rast_id, field = "gridcode")

## plot to see what it looks like
plot(bathy_rast)
```

#### **STEP 4: Reclassify the depth classes so only show the Upper Shelf (~50m depth)**
```{r}
## create a matrix of needed classes
m <- c(0.00, 7, NA, # made NA (instead of 0) for distance decay 
       8, 8, 1)

### wrap the vector into a matrix by rows
rcl_mtx <- matrix(m, ncol=3, byrow=TRUE)

## reclassify with the new matrix
bathy_reclass <- reclassify(bathy_rast, rcl_mtx)

## plot it to check
plot(bathy_reclass)

### save it
writeRaster(bathy_reclass, here('set_up/final_rasters/costs/bathy_rast.tif'), overwrite = TRUE)
```


#### **STEP 5: Distance Decay with 50m priorization (from above)**
```{r}
## use the raster::distance() function, which calculates a distance for every NA cell to the nearest non-NA cell.
dist_rast <- raster::distance(bathy_reclass) %>% 
  mask(mz_eez_sf) # mask so the land distances are not calculated

## plot it!
plot(dist_rast)

## change 0 values to 0.00001 since prioritzr 
dist_rast[dist_rast == 0] <-0.00001

## plot again 
plot(dist_rast)

## rescale from 0-1
dist_rescaled <- dist_rast/maxValue(dist_rast)

## prioritzr doesn't like values lower than 1 x10^-6 so we made it 0.00001
dist_rescaled[dist_rescaled <= 2.789642e-10] <- 0.00001

## plot it
plot(dist_rescaled)

## created a dataframe to check it out
dist_df <- dist_rast %>%
  rasterToPoints() %>% 
  as.data.frame()

dist_rescaled_df <- dist_rescaled %>% 
  rasterToPoints() %>% 
  as.data.frame()


### save it
writeRaster(dist_rescaled, here('set_up/final_rasters/costs/dist_50_rast.tif'), overwrite = TRUE)

```

 
Outputs:

 - `bathy_rast.tif` = raster of bathymetry in Mozambique (<50m depth)
 - `dist_50_rast.tif` = distance decay raster of bathymetry in Mozambique (<50m depth)

End script.



