---
title: "species_status"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(fasterize)
```

# Extra Cost: Species Status
This cost layer looks at prioritizing species with more at risk conservation status. This was just an exploration and was not used in our final analysis.

#### **STEP 1: Read in Mozambique EEZ and Shapefile**
```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))
plot(mz_rast_id)

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

```

#### **STEP 2: Read in IUCN CHONDRICHTHYES Shapefiles**
```{r}
### grab iucn data from local computer
iucn_path <- 'G:/group_project/data/updated_iucn'
iucn_shape <- list.files(iucn_path, full.names = TRUE) 

iucn_sf <- sf::read_sf(iucn_shape[9])

### create a dataframe to look at the different variables and remove geometry to speed it up
iucn_df <- iucn_sf %>%
  as.data.frame() %>% 
  select(-geometry)

```


#### **STEP 3: Reclassifying based on conservation category/status**
```{r}
### reclassify status
risk_df <- iucn_sf  %>% 
  mutate(risk_level = case_when(category == "CR" ~3,
                          category == "EN" ~2,
                          category == "VU"~ 1))

### check to see if it works with thresher sharks
### create dataframe that only contains species and changes to correct crs
thresher_shark <- risk_df %>%
  filter(binomial == "Alopias pelagicus") %>% 
  st_transform(., crs = st_crs(mz_eez_sf)) 

### check crs to make sure it matches
st_crs(thresher_shark)

### create a raster
thresher_shark_rast <- fasterize::fasterize(thresher_shark, mz_rast_id, field = "risk_level") %>%
  mask(mz_eez_sf)

### plot to see how it looks
plot(thresher_shark_rast)

```

#### **STEP 4: Create a Function that Creates a Raster for Each Species**
We created a function that based on the steps above (species sf --> species raster) and saves the raster in our designated folder for finalized species rasters using risk level instead 
```{r}
create_rast_status <- function(risk_df, species_name){
  outfile <- sprintf(here('set_up/final_rasters/conservation_features/iucn_species_status/species_%s.tif'), species_name)
  message('Processing species ', species_name, ' to create file ', outfile)

  species_df <- risk_df %>%
  filter(binomial == species_name) %>% 
  st_transform(., crs = st_crs(mz_eez_sf))

species_rast <- fasterize::fasterize(species_df, mz_rast_id, field = "risk_level") %>% 
  mask(mz_eez_sf)

writeRaster(species_rast, filename = outfile, overwrite = TRUE)

return('yay it works')

}

# dusky_shark_rast <- create_rast_status(risk_df, "Carcharhinus obscurus")
# plot(dusky_shark_rast)

```

#### **STEP 5: Create Vector of Species Names for Loop**
This vector contains our focal shark and ray species by scientific name. You can add more or less species to this vector.
```{r}
### create vector 
species_name <- c("Alopias pelagicus",
                 "Alopias superciliosus",
                 "Alopias vulpinus",
                 "Carcharhinus amblyrhynchos",
                 "Carcharhinus longimanus",
                 "Carcharhinus obscurus",
                 "Carcharodon carcharias",
                 "Isurus oxyrinchus",
                 "Isurus paucus",
                 "Carcharias taurus",
                 "Rhincodon typus",
                 "Sphyrna lewini",
                 "Sphyrna mokarran",
                 "Sphyrna zygaena",
                 "Stegostoma tigrinum",
                 "Mobula alfredi",
                 "Mobula birostris",
                 "Mobula kuhlii",
                 "Mobula mobular",
                 "Pristis pristis",
                 "Pristis zijsron",
                 "Rostroraja alba",
                 "Rhina ancylostoma",
                 "Acroteriobatus leucospilus")
```

#### **STEP 6: Create & Run Loop with Species Vector**
```{r}
### create loop 
for (i in seq_along(species_name)) {
  create_rast_status(risk_df, species_name[i])
}
```

#### **STEP 7: Creating Species Stack** --must clear environment to run this chuck...
```{r}
### create file path to read in all 27 species tifs in 
species_status_path <- here("set_up", "final_rasters", "conservation_features", "iucn_species_status")
species_status_files <- list.files(species_status_path, full.names = TRUE)

### stack them (this includes the three that we manually created and saved in the same IUCN folder)
species_status_stack <- stack(species_status_files)

### plot to see how they look and make a data frame to double check
plot(species_status_stack)
species_status_stack_df <- as.data.frame(species_status_stack)

### write raster
writeRaster(species_status_stack, here('set_up/final_rasters/conservation_features/iucn_species_status/species_status_stack.tif'), options="INTERLEAVE=BAND",  overwrite = TRUE)
```


Citations:
 - IUCN (International Union for Conservation of Nature) 2021. *Chondrichthyes (sharks, rays and chimaeras). The IUCN Red List of Threatened Species*. Version 6.2. https://www.iucnredlist.org. Downloaded on 24 Janurary 2021.
 
Outputs:
 - `species_XXX XXX` = tif files of each focal species to be used with risk level as weight (27 total)
 - `species_status_stack.tif` = raster stack of all focal species for conservation features with risk level
 
 
End script.


