---
title: "1_MZ_EEZ_raster"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages 
#Since raster has a "select" function but we more often use the tidyverse::select, load raster first, and let tidyverse overwrite that function.
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)

```

# Establishing the Project Area

Before we can begin our analysis, we need to establish the project area which in this case is Mozambique's Exclusive Economic Zone (EEZ). This shapefile was provided by our client Wildlife Conservation Society (WCS). Similar steps can be followed even with a different project area if initial dataset is a shapefile. 

#### **STEP 1: Setting up the MZ EEZ File**
```{r, include=FALSE}

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

## Read in shapefile as simple feature 
mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## Add a buffer of 10 km to make sure cells on the outer edges of the EEZ are included
mz_eez_buffer_sf <- sf::st_buffer(mz_eez_sf, dist = 10000)

## Create data frame from simple feature to see data more readily (not necessary, but helpful for larger sf)
mz_eez_df <- mz_eez_sf %>% as.data.frame() %>% select(-geometry)

## Use Moz EEZ to set extent for future rasters
mz_ext <- raster::extent(mz_eez_buffer_sf)
# class      : Extent 
# xmin       : 3119112 
# xmax       : 4159767 
# ymin       : -3412817 
# ymax       : -1270921 

```

#### **STEP 2: Create MZ raster with Cell IDs**
In order to rasterize other shapefiles we will be using in the future, we assigned cell values to our project area raster.

```{r, include=FALSE}
## Create raster with cell ids and clip to MZ EEZ raster
mz_rast_id <- raster::raster(x=mz_ext, crs=crs(mz_eez_sf), res=10000)

## Assign cell values
values(mz_rast_id) <- 1:ncell(mz_rast_id) 

## Create data frame to check it out
mz_rast_id_df <- rasterToPoints(mz_rast_id) %>% as.data.frame()

## Plot to make sure it looks good
plot(mz_rast_id)

## Mask it to only include the EEZ
mz_rast_id <- mask(mz_rast_id, mz_eez_sf)
plot(mz_rast_id)

## Let's save this cell id raster as tif file - hashtagged this out for now since the tif is too big
writeRaster(mz_rast_id, here('set_up/final_rasters/mz_eez_templates/mz_rast_id.tif'), overwrite = TRUE)

```

#### **STEP 3: Create raster with higher res of 1000 m**
Also, we need to set up rast id with higher resolution of 1000 m to use when we rasterize some of our conservation feature layers that are at a finer resolution like coral and mangroves

```{r, include=FALSE}

## Create raster with res=100
#mz_eez_rast_100 <- raster::raster(x=extent(mz_rast_id), crs=crs(mz_eez_sf), res=100)

## Assign cell values
#values(mz_eez_rast_100) <- 1:ncell(mz_eez_rast_100)

## Mask to Moz EEZ
#mz_eez_rast_100 <- mask(mz_eez_rast_100, mz_eez_buffer_sf)

## Plot and create data frame to validate it looks good
#plot(mz_eez_rast_100)
#mz_rast_id_df <- rasterToPoints(mz_eez_rast_100) %>% as.data.frame()


### Use Moz EEZ to raster with 1000 m resolution 
mz_eez_rast_1000 <- raster::raster(x=extent(mz_rast_id), crs=crs(mz_eez_sf), res=1000)

## Assign cell values
values(mz_eez_rast_1000) <- 1:ncell(mz_eez_rast_1000)

## Mask to Moz EEZ
mz_eez_rast_1000 <- mask(mz_eez_rast_1000, mz_eez_sf)

## Plot and create data frame to validate it looks good
plot(mz_eez_rast_1000)
mz_rast_id_df <- rasterToPoints(mz_eez_rast_1000) %>% as.data.frame()

## Let's save this mz raster with cell ids as tif file
#writeRaster(mz_eez_rast_100, here('set_up/final_rasters/mz_eez_templates/mz_eez_rast_100.tif'), overwrite = TRUE)

## Let's save this mz raster with cell ids as tif file
writeRaster(mz_eez_rast_1000, here('set_up/final_rasters/mz_eez_templates/mz_eez_rast_1000.tif'), overwrite = TRUE)

```

Outputs:

- `mz_rast_id` = raster of the Mozambique EEZ at 10,000 resolution with cell ids
- `mz_rast_100` = raster of the Mozambique EEZ at 100 resolution with cell ids
- `mz_rast_1000`= raster of the Mozambique EEZ at 1,000 resolution with cell ids
- `mz_eez_sf` = simple features of the Mozambique EEZ

End script.
