---
---
title: "1_MZ_EEZ_raster"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages 
#Since raster has a "select" function but we more often use the tidyverse::select, load raster first, and let tidyverse overwrite that function.
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)

```

## *STEP 1: Setting up the MZ EEZ File*
```{r, include=FALSE}

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

## Read in shapefile as simple feature 
mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## Add a buffer of 10 km to make sure cells on the outer edges of the EEZ are included
mz_eez_buffer_sf <- sf::st_buffer(mz_eez_sf, dist = 10000)

## Create data frame from simple feature to see data more readily (not necessary, but helpful for larger sf)
mz_eez_df <- mz_eez_sf %>% as.data.frame() %>% select(-geometry)

## Use Moz EEZ to set extent for future rasters
mz_ext <- raster::extent(mz_eez_buffer_sf)
# class      : Extent 
# xmin       : 3119112 
# xmax       : 4159767 
# ymin       : -3412817 
# ymax       : -1270921 

```

### *Create MZ raster with Cell IDs*
```{r, include=FALSE}

## Create raster with cell ids and clip to MZ EEZ raster
mz_rast_id <- raster::raster(x=mz_ext, crs=crs(mz_eez_sf), res=10000)

## Assign cell values
values(mz_rast_id) <- 1:ncell(mz_rast_id) 

## Create data frame to check it out
mz_rast_id_df <- rasterToPoints(mz_rast_id) %>% as.data.frame()

## Plot to make sure it looks good
plot(mz_rast_id)

## Mask it to only include the EEZ
mz_rast_id <- mask(mz_rast_id, mz_eez_sf)
plot(mz_rast_id)

## Let's save this cell id raster as tif file - hashtagged this out for now since the tif is too big
writeRaster(mz_rast_id, here('set_up/final_rasters/mz_eez_templates/mz_rast_id.tif'), overwrite = TRUE)

### Let's save this mz raster with cell ids as tif file
#writeRaster(mz_eez_rast, here('set_up/final_rasters/mz_eez_templates/mz_eez_rast.tif'), overwrite = TRUE) --> do we need this?

```


### *Create raster with higher res of 1000 m*
Also need to set up rast id with higher resolution of 1000 m to use when we rasterize some of our conservation feature layers that are at a finer resolution like coral and mangroves
```{r, include=FALSE}

## Create raster with res=100 
#mz_eez_rast_100 <- raster::raster(x=extent(mz_rast_id), crs=crs(mz_eez_sf), res=100)
 
## Assign cell values
#values(mz_eez_rast_100) <- 1:ncell(mz_eez_rast_100)

## Mask to Moz EEZ
#mz_eez_rast_100 <- mask(mz_eez_rast_100, mz_eez_buffer_sf)

## Plot and create data frame to validate it looks good
#plot(mz_eez_rast_100)
#mz_rast_id_df <- rasterToPoints(mz_eez_rast_100) %>% as.data.frame()


### Use Moz EEZ to raster with 1000 m resolution 
mz_eez_rast_1000 <- raster::raster(x=extent(mz_rast_id), crs=crs(mz_eez_sf), res=1000)

## Assign cell values
values(mz_eez_rast_1000) <- 1:ncell(mz_eez_rast_1000)

## Mask to Moz EEZ
mz_eez_rast_1000 <- mask(mz_eez_rast_1000, mz_eez_sf)

## Plot and create data frame to validate it looks good
plot(mz_eez_rast_1000)
mz_rast_id_df <- rasterToPoints(mz_eez_rast_1000) %>% as.data.frame()


## Let's save this mz raster with cell ids as tif file
#writeRaster(mz_eez_rast_100, here('set_up/final_rasters/mz_eez_templates/mz_eez_rast_100.tif'), overwrite = TRUE)

## Let's save this mz raster with cell ids as tif file
writeRaster(mz_eez_rast_1000, here('set_up/final_rasters/mz_eez_templates/mz_eez_rast_1000.tif'), overwrite = TRUE)

```

Citations:
 - 
 
Outputs:

- `mz_rast_id` = raster of the Mozambique EEZ at 10000 resolution with cell ids
- `mz_rast_100` = raster of the Mozambique EEZ at 100 resolution with cell ids
- `mz_rast_1000`= raster of the Mozambique EEZ at 1000 resolution with cell ids
= `mz_eez_sf` = simple features of the Mozambique EEZ

End script.
