---
title: "2_iucn_species"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(stringr)
library(janitor)
library(fasterize)
```

#### **STEP 1: Read in Mozambique EEZ and Shapefile**
```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))
plot(mz_rast_id)

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## Add a buffer of 10 km to make sure cells on the outer edges of the EEZ are included
#mz_eez_buffer_sf <- sf::st_buffer(mz_eez_sf, dist = 10000)
```


#### **STEP 2: Read in IUCN CHONDRICHTHYES Shapefiles**
```{r}
### grab iucn data from local computer
iucn_path <- 'G:/group_project/data/updated_iucn'
iucn_shape <- list.files(iucn_path, full.names = TRUE) 

iucn_sf <- sf::read_sf(iucn_shape[9])

### create a dataframe to look at the different variables and remove geometry to speed it up
iucn_df <- iucn_sf %>%
  as.data.frame() %>% 
  select(-geometry)

```


#### **STEP 3: Test with Two Species**
```{r}
### create dataframe that only contains species and changes to correct crs
thresher_shark <- iucn_sf %>% 
  filter(binomial == "Alopias pelagicus") %>% 
  st_transform(., crs = st_crs(mz_eez_sf))

### check crs to make sure it matches
st_crs(thresher_shark)

### create a raster 
thresher_shark_rast <- fasterize::fasterize(thresher_shark, mz_rast_id) %>% 
  mask(mz_eez_sf)

### plot to see how it looks
plot(thresher_shark_rast)

### made all NA = 0 (not sure if this makes a difference)
thresher_shark_rast[is.na(thresher_shark_rast[])] <- 0
plot(thresher_shark_rast)

### masked it to make all values outside the eez = NA
thresher_shark_rast <- mask(thresher_shark_rast, mz_eez_sf)
plot(thresher_shark_rast)

### create dataframe that only contains species and changes to correct crs
dusky_shark <- iucn_sf %>% 
  filter(binomial == "Carcharhinus obscurus") %>% 
  st_transform(., crs = st_crs(mz_eez_sf))

### check crs to make sure it matches
st_crs(dusky_shark)

### create a raster 
dusky_shark_rast <- fasterize::fasterize(dusky_shark, mz_rast_id) %>% 
  mask(mz_eez_sf)

### plot to see how it looks
plot(dusky_shark_rast)

### create dataframe that only contains species and changes to correct crs
shark <- iucn_sf %>% 
  filter(binomial == "Carcharhinus amblyrhynchos") %>% 
  st_transform(., crs = st_crs(mz_eez_sf))

### check crs to make sure it matches
st_crs(shark)

### create a raster 
shark_rast <- fasterize::fasterize(shark, mz_rast_id) %>% 
  mask(mz_eez_sf)

### plot to see how it looks
plot(shark_rast)
```

#### **STEP 4: Create a Function that Creates a Raster for Each Species**
```{r}
# #create raster function by species name
create_rast <- function(iucn_sf, species_name){
  outfile <- sprintf(here('set_up/final_rasters/conservation_features/iucn_species/species_%s.tif'), species_name)
  message('Processing species ', species_name, ' to create file ', outfile)

  species_df <- iucn_sf %>%
  filter(binomial == species_name) %>% 
  st_transform(., crs = st_crs(mz_eez_sf))

species_rast <- fasterize::fasterize(species_df, mz_rast_id) %>% 
  mask(mz_eez_sf)

writeRaster(species_rast, filename = outfile, overwrite = TRUE)

return('yay it works')

}
```

#### **STEP 5: Create Vector of Species Names for Loop**
```{r}
### create vector 
species_name <- c("Alopias pelagicus",
                 "Alopias superciliosus",
                 "Alopias vulpinus",
                 "Carcharhinus amblyrhynchos",
                 "Carcharhinus longimanus",
                 "Carcharhinus obscurus",
                 "Carcharodon carcharias",
                 "Isurus oxyrinchus",
                 "Isurus paucus",
                 "Carcharias taurus",
                 "Rhincodon typus",
                 "Sphyrna lewini",
                 "Sphyrna mokarran",
                 "Sphyrna zygaena",
                 "Stegostoma tigrinum",
                 "Mobula alfredi",
                 "Mobula birostris",
                 "Mobula kuhlii",
                 "Mobula mobular",
                 "Pristis pristis",
                 "Pristis zijsron",
                 "Rostroraja alba",
                 "Rhina ancylostoma",
                 "Acroteriobatus leucospilus")
                 
```

#### **STEP 6: Create & Run Loop with Species Vector**
```{r}
### create loop 
for (i in seq_along(species_name)) {
  create_rast(iucn_sf, species_name[i])
}
```

#### **STEP 7: Pull in the three IUCN distributions that were refined with WCS experts**
```{r}
### read in the range map for Pseudoginglymostoma brevicaudatum:
p_brevicaudatum <-here("set_up", "raw_data", "pseudoginglymostoma_brevicaudatum")
p_brevicaudatum_shape <- list.files(p_brevicaudatum, full.names = TRUE) 

p_brevicaudatum_sf <- sf::read_sf(p_brevicaudatum_shape[7]) %>% 
    st_transform(., crs = st_crs(mz_eez_sf))

p_brevicaudatum_rast <- fasterize::fasterize(p_brevicaudatum_sf, mz_rast_id) %>% 
  mask(mz_eez_sf)

plot(p_brevicaudatum_rast)

writeRaster(p_brevicaudatum_rast, here('set_up/final_rasters/conservation_features/iucn_species/species_Pseudoginglymostoma_brevicaudatum_wcs.tif'), overwrite = TRUE)

### read in the range map for Rhynchobatus australiae:
r_ausrtraliae <-here("set_up", "raw_data", "rhynchobatus_australiae")
r_ausrtraliae_shape <- list.files(r_ausrtraliae, full.names = TRUE) 

r_australiae_sf <- sf::read_sf(r_ausrtraliae_shape[7]) %>% 
    st_transform(., crs = st_crs(mz_eez_sf))

r_australiae_rast <- fasterize::fasterize(r_australiae_sf, mz_rast_id) %>% 
  mask(mz_eez_sf)

plot(r_australiae_rast)

writeRaster(r_australiae_rast, here('set_up/final_rasters/conservation_features/iucn_species/species_Rhynchobatus_australiae_wcs.tif'), overwrite = TRUE)

### read in the range map for Rhynchobatus djiddensis:
r_djiddensis <-here("set_up", "raw_data", "rhynchobatus_djiddensis")
r_djiddensis_shape <- list.files(r_djiddensis, full.names = TRUE) 

r_djiddensis_sf <- sf::read_sf(r_djiddensis_shape[7]) %>% 
    st_transform(., crs = st_crs(mz_eez_sf))

r_djiddensis_rast <- fasterize::fasterize(r_djiddensis_sf, mz_rast_id) %>% 
  mask(mz_eez_sf)

plot(r_djiddensis_rast)

writeRaster(r_djiddensis_rast, here('set_up/final_rasters/conservation_features/iucn_species/species_Rhynchobatus_djiddensis_wcs.tif'), overwrite = TRUE)

```


#### **STEP 8: Creating Species Stack** --must clear environment to run this chuck...
```{r}
### create file path to read in all 30 species tifs in 
species_path <- here("set_up", "final_rasters", "conservation_features", "iucn_species")
species_files <- list.files(species_path, full.names = TRUE)

### stack them (this includes the three that we manually created and saved in the same IUCN folder)
species_stack <- stack(species_files)

### plot to see how they look and make a data frame to double check
plot(species_stack)
species_stack_df <- as.data.frame(species_stack)

### write raster
writeRaster(species_stack, here('set_up/final_rasters/conservation_features/species_stack.tif'), options="INTERLEAVE=BAND",  overwrite = TRUE)
```

Citations:
 - 
 
Outputs:
 - `species_XXX XXX` = tif files of each focal species to be used (27 total)
 - `species_stack.tif` = raster stack of all focal species for conservation features
 
End script.



