---
title: "Conservation Features Set Up"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load in packages
library(raster)
library(tidyverse)
library(sf)
library(here)
library(fasterize)
library(janitor)
```

### **STEP 1: Read in Mozambique EEZ and Shapefile**
```{r}
## read in MZ EEZ tif
mz_rast_id <- raster(here("set_up/final_rasters/mz_eez_templates/mz_rast_id.tif"))

## Grab Moz Exclusive Economic Zone (EEZ) shapefile from local computer
mz_eez_path <- 'G:/group_project/data/mz_eez'
mz_shape <- list.files(mz_eez_path, full.names = TRUE) 

mz_eez_sf <- sf::read_sf(mz_shape[6])

## check coordinate system
st_crs(mz_eez_sf)

## Add a buffer of 10 km to make sure cells on the outer edges of the EEZ are included
mz_eez_buffer_sf <- sf::st_buffer(mz_eez_sf, dist = 10000)

## Grab Moz Exclusive Economic Zone (EEZ)  at 100m shapefile from local computer
mz_eez_100_path <- 'G:/group_project/data/mz_raster'
mz_100_shape <- list.files(mz_eez_100_path, full.names = TRUE) 

mz_rast_100 <- raster(mz_100_shape[1])

```

### **STEP 2: Read in ACA data for Coral and Seagrass Shapefiles**
```{r}
## Grab aca shapefile from local computer
## Path to get data on G shared drive at Bren
aca_path <- 'G:/group_project/Data/Habitats/ACA_classes'
aca_shape <- list.files(aca_path, full.names = TRUE) 

## Read in shapefile as a simple feature and tranform to the correct crs
aca_sf <- sf::read_sf(aca_shape[8])%>% 
  st_transform(., crs = st_crs(mz_eez_sf))

## Check the CRS of the coral sf to make sure it matches the CRS of mz_rast
crs(mz_rast_id)
st_crs(aca_sf)

## Create data frame to easily check data inside sf
aca_df <- aca_sf %>% 
  as.data.frame() %>% 
  select(-geometry) 
```

### **STEP 3: Create all habitat rasters**

#### CORAL
```{r}
### coral shapefile
aca_coral_sf <- aca_sf %>% 
  filter(Ecosystem == "Coral Reefs") %>% 
  select(-class)

## Make a raster of the coral simple feature using mz_rast_100 - the raster we created at a smaller resolution to set extent and mask to the mz_eez_buffer_sf
aca_coral_raster <- fasterize::fasterize(aca_coral_sf, mz_rast_100)%>%
  mask(mz_eez_sf)

## plot to make sure it looks good
plot(aca_coral_raster) # look closely it's there!

## Save as raster
writeRaster(aca_coral_raster, here('set_up/final_rasters/conservation_features/coral/aca_coral.tif'), overwrite = TRUE)

# Aggregate to become a coarser resolution at 10000 m
aca_coral_raster_agg <- raster::aggregate(aca_coral_raster, fact=100, fun=sum, na.rm = TRUE)
aca_coral_aggregate_per <- aca_coral_raster_agg/10000

## mask to the MZ EEZ
aca_coral_aggregate_per <- mask(aca_coral_aggregate_per, mz_eez_sf)

## Lets plot to see what it looks like
plot(aca_coral_aggregate_per)

## Save coral aggregate as raster
writeRaster(aca_coral_aggregate_per, here('set_up/final_rasters/conservation_features/coral/aca_coral_agg.tif'), overwrite = TRUE)
```


#### SEAGRASS
```{r}
### seagrass shapefile
aca_seagrass_sf <- aca_sf %>% 
  filter(Ecosystem == "Seagrass") %>% 
  select(-class)

## Let's create a rsater using the mz_rast with resolution of 10000 and mask to the MZ EEZ
aca_seagrass_raster <- fasterize::fasterize(aca_seagrass_sf, mz_rast_100) %>% 
  mask(mz_eez_sf)

## plot to make sure it looks good
plot(aca_seagrass_raster)

## Save as raster
writeRaster(aca_seagrass_raster, here('set_up/final_rasters/conservation_features/seagrass/aca_seagrass.tif'), overwrite = TRUE)

# Aggregate to become a coarser resolution at 10000 m
aca_seagrass_raster_agg <- raster::aggregate(aca_seagrass_raster, fact=100, fun=sum, na.rm = TRUE)
aca_seagrass_aggregate_per <- aca_seagrass_raster_agg/10000

aca_seagrass_aggregate_per <- mask(aca_seagrass_aggregate_per, mz_eez_sf)

plot(aca_seagrass_aggregate_per)

## Save coral aggregate as raster
writeRaster(aca_seagrass_aggregate_per, here('set_up/final_rasters/conservation_features/seagrass/aca_seagrass_agg.tif'), overwrite = TRUE)
```


#### MANGROVES
```{r}
## Path to get data on G shared drive at Bren
mangrove_path <- 'G:/group_project/Data/Habitats/mangroves'
mangrove_shape <- list.files(mangrove_path, full.names = TRUE)

## Read in shapefile as a simple feature and select relevant columns
mangrove_sf <- sf::read_sf(mangrove_shape[6]) %>%
  select("OBJECTID", "ISO3", "AREA_KM2", "AREA_M2", "geometry")

## Plot to see what simple feature looks like
plot(mangrove_sf)

## Check the CRS of the mangrove sf and compare to mz_rast
st_crs(mz_rast_id)
st_crs(mangrove_sf)

## Create data frame to easily check data inside coral sf
mangrove_df <- mangrove_sf %>% as.data.frame() %>% select(-geometry)

## Mangroves also has finer resolution that we want to be sure we capture so lets make a raster of the mangove simple feature using mz_rast_100 - the raster we created at a smaller resolution to set extent and mask to the mz_eez
mangrove_raster <- fasterize::fasterize(mangrove_sf, mz_rast_100)%>%
  mask(mz_eez_sf)

## plot to make sure it looks good
plot(mangrove_raster)

## Now we will aggregate to become a coaser resoltuion that matches our other rasters at 10000 m
mangrove_agg <- raster::aggregate(mangrove_raster, fact=100, fun=sum, na.rm=TRUE)
mangrove_agg_percent <- mangrove_agg/10000

mangrove_rast <- mask(mangrove_agg_percent, mz_eez_sf)

## Lets plot to see what it looks like
plot(mangrove_rast)

## Lookgs good - let's save as a tif file
writeRaster(mangrove_agg_percent, here('set_up/final_rasters/conservation_features/mangrove/mangrove.tif'), overwrite = TRUE)

```

#### SEAMOUNTS
```{r}
## Grab seamount shapefile from the local computer
seamount_path <- 'G:/group_project/Data/Habitats/seamounts'
seamount_shape <- list.files(seamount_path, full.names = TRUE)

## Load shapefile as simple feature
seamount_sf <- sf::read_sf(seamount_shape[6]) 

## Check the CRS of the seamount and compare to the mz rast
st_crs(mz_rast_id)
st_crs(seamount_sf)

## Let's create a rsater using the mz_rast with resolution of 10000 and mask tot the MZ EEZ
seamount_raster <- fasterize::fasterize(seamount_sf, mz_rast_id)%>%
  mask(mz_eez_sf)

## Plot to see what it looks like
plot(seamount_raster)

# Looks good lets save as a tif
writeRaster(seamount_raster, here('set_up/final_rasters/conservation_features/seamounts/seamount.tif'), overwrite = TRUE)
```

#### KNOLLS
```{r}
## Grab knolls shapefile from the local computer
knolls_path <- 'G:/group_project/Data/Habitats/knolls'
knolls_shape <- list.files(knolls_path, full.names = TRUE)

## Load shapefile as simple feature
knolls_sf <- sf::read_sf(knolls_shape[6]) 

## Check the CRS of the knolls sf
st_crs(knolls_sf)

## Let's create a rsater using the mz_rast with resolution of 10000 and mask tot the MZ EEZ
knolls_raster <- fasterize::fasterize(knolls_sf, mz_rast_id)%>%
  mask(mz_eez_sf)

## Plot to see what it looks like
plot(knolls_raster)

# Looks good - lets save as a tif
writeRaster(knolls_raster, here('set_up/final_rasters/conservation_features/knolls/knolls.tif'), overwrite = TRUE)

```
### **STEP 3: Create raster stack**

```{r}
## Create a features stack from all five critical habitats created in the previous steps
habitat_stack <- stack(aca_coral_raster_agg, mangrove_rast, aca_seagrass_raster_agg, seamount_raster, knolls_raster)

## Plot to see what it looks like
plot(habitat_stack)

plot(habitat_stack, main = paste("Feature", seq_len(nlayers(habitat_stack))),
     nr = 2, box = FALSE, axes = FALSE)

## Looks good, let's save as a multilayer feature tif
writeRaster(habitat_stack, filename=here("set_up","final_rasters","conservation_features", "habitat_stack.tif"), options="INTERLEAVE=BAND", overwrite=TRUE)
```

### **STEP 4: Combine habitat raster stack and species raster stack to create stack of all conservation features**
```{r}
## pull in species stack we created in step 2
species_stack <- brick(here("set_up", "final_rasters", "conservation_features", "species_stack.tif"))

plot(species_stack)

### stack all species and habitat
features_stack <- stack(habitat_stack, species_stack)

plot(features_stack)

### change into a dataframe to check it
features_df <- as.data.frame(features_stack)

### write raster
writeRaster(features_stack, here('set_up/final_rasters/conservation_features/all_features_stack.tif'),options="INTERLEAVE=BAND", overwrite = TRUE)
```



Citations:
 - 
 
Outputs:
 -`aca_coral` = raster of coral within Mozambique EEZ at 100m resolution
 -`aca_coral_agg` = raster of coral within Mozambique EEZ aggregated to match 10000m resolution
 -`aca_seagrass` = raster of seagrass within Mozambique EEZ at 100m resolution
 -`aca_seagrass_agg` = raster of seagrass within Mozambique EEZ aggregated to match 10000m resolution
 -`mangrove` = raster of mangrove within Mozambique EEZ at 100m resolution
 -`mangrove_agg` = raster of mangrove within Mozambique EEZ aggregated to match 10000m resolution
 -`knolls` = raster of knolls within Mozambique EEZ at 10000m resolution
 -`seamounts` = raster of seamounts within Mozambique EEZ at 10000m resolution
 -`habitat_stack` = raster stack of all critical habitats

 
End script.







